---
title: "Heartbeat evoked Potential in FND"
author: "Natascha Stoffel"
date: "`r date()`"
format:
  html:
    embed-resources: true
    toc: true #to show table of contents
editor: visual
---

# Information on the Data

BASEC registration number : 2023-0469

Study Visit "Interoception" conducted between **18.10.2023 - 4.9.2024**, with completed **N=92 participants.**

Analysis conducted with RStudio 2023.12.1+402.

# 0) Preparation

```{r}
#| output: false
#| warning: FALSE
#| error: FALSE
# remove everything that we stored in the environment
rm(list=ls()) 

# Load necessary libraries
library(readr)      # For read_csv()
library(dplyr)      # For data manipulation functions
library(gtsummary)  # For tbl_summary()
library(gt)         # For working with gt tables
library(ggplot2)    # For plotting
library(ggpubr)     # For ggboxplot() and adding jitter
library(ggsignif)   # For statistical comparisons in plots
library(gridExtra)  # For arranging multiple plots in a grid
library(openxlsx)   # For saving Excel files
library(tidyr)      # For reshaping data
library(plyr)       # For revalue() function
library(car)        # testing assumptions in models
library(rstatix)    # running ancova models
library(lmerTest)   # running linear mixed models
library(writexl)    # save excel files
library(ARTool)     # For ANOVA of non-normal distribution

# load prepared and cleaned file from 00_preparation: df_yyyymmdd.csv
setwd("/Users/nataschastoffel/Documents/GitHub/interoception_NS")
Data.CRF <- read.xlsx("~/Documents/GitHub/interoception_NS/data/processed/12-2024/Data.CRF_FINAL_20241218.xlsx")

## LOAD  EEG DATA  -----------------------------------------------------------
# load prepared and cleaned file from 00_preparation: df_yyyymmdd.csv
df_EEG <- read.xlsx("/Users/nataschastoffel/Documents/GitHub/interoception_NS/data/processed/02-2025/AverageTimeInterval_HEP_summary.xlsx")

df_EEG$pcode <- substr(df_EEG$pcode, 1, 4) # keeps only the first four digits (P001 etc) in the names
df_EEG$pcode <- sub("^P00", "", df_EEG$pcode) # gets rid of the P00 for all participants
df_EEG$pcode <- sub("^P0", "", df_EEG$pcode) # gets rid of the P0 for all participants
df_EEG$pcode <- sub("^P", "", df_EEG$pcode) # gets rid of the P for all participants

df<-merge(Data.CRF, df_EEG, x.by=c("pcode"),y.by=c("pcode"),all.x=TRUE, all.y=TRUE)
# View(df)

df <- df[complete.cases(df$F8_HEART_510), ] # chose any of the eeg variables; so we remove those participants in our df, that did not complete the HEP study
# N = 88 in EEG sample

## uploade ECG amplitude (normalized traces)
df_ecg_amp <- read.xlsx("/Users/nataschastoffel/Documents/GitHub/interoception_NS/data/processed/02-2025/ECG_normalized_amplitude.xlsx")
df_ecg_amp$pcode <- sub("^P00", "", df_ecg_amp$pcode) # gets rid of the P00 for all participants
df_ecg_amp$pcode <- sub("^P0", "", df_ecg_amp$pcode) # gets rid of the P0 for all participants
df_ecg_amp$pcode <- sub("^P", "", df_ecg_amp$pcode) # gets rid of the P for all participants

df <- df %>% mutate(pcode = as.character(pcode))
df_ecg_amp <- df_ecg_amp %>% mutate(pcode = as.character(pcode))
df<-merge(df, df_ecg_amp, x.by=c("pcode"),y.by=c("pcode"),all.x=TRUE, all.y=TRUE)
# View(df)

# define medication intake of interest; those with a considerable effect on the BRAIN (which is all psychotropic medication, benzos, morphiom/opiod pain killers and antiepileitc medication)
df <- df %>%
  mutate(medication_brain = if_else(psychotropic_medication == 1 | sleep_benzo == 1 | pain_antiepi == 1 | pain_morph_opio == 1, 1, 0))  # 1=yes so if there is a yes in any of the different medication categories, there will be a 1 in medication_brain
df %>% filter(is.na(medication_brain)) %>%  select(pcode) # P057 has a NA; manually check, and complement with answer =  NO medication 
df <- df %>%  mutate(medication_brain = ifelse(pcode == 57, 0, medication_brain))

# create a common depression-anxiety variable as it highly correlates (and to not mess with the models)
df <- df %>%
  mutate(anx_dep_SUM = stai_t + bdi) # create one variable for the trait anxiety and depression to use as one variable in the model

#making sure everything is stored as the correct variable
df$group <- as.factor(df$group)  
df$sex <- as.factor(df$sex)      
df$medication_brain <- as.factor(df$medication_brain)  
df$menopause <- as.factor(df$menopause)  
df$contraception <- as.factor(df$contraception)  
df$mc_regularity <- as.factor(df$mc_regularity)  

# define required subgroups
df_female <- df[df$sex == "1", ]
df_male <- df[df$sex == "2", ]

df_HC <- df[df$group == "0", ]
df_FND <- df[df$group == "1", ]

df_med <- df[df$medication_brain == "1", ]
df_med <- df_med[complete.cases(df_med$pcode), ] 
df_nomed <- df[df$medication_brain == "0", ]
df_nomed <- df_nomed[complete.cases(df_nomed$pcode), ] 

```

# 1) Demographics

# 1.1 Difference between groups

```{r}
#| warning: FALSE
#| error: FALSE
# AGE
shapiro.test(df$age) #  normally distributed with  p < 0.05
res<-t.test(df$age[df$group=="1"], df$age[df$group=="0"])
res # p-value = 0.6338
summary(df_FND$age)
summary(df_HC$age)

## check subsample of non-medicated
  shapiro.test(df_nomed$age) #  normally distributed with  p < 0.05
  res<-t.test(df_nomed$age[df_nomed$group=="1"], df_nomed$age[df_nomed$group=="0"])
  res # p = 0.06917
  summary(df_nomed$age[df_nomed$group=="0"])
  summary(df_nomed$age[df_nomed$group=="1"])

# SEX
table_sex_group <- table(df$sex, df$group)
table_sex_group
summary(df_FND$sex)
summary(df_HC$sex)
# Perform the chi-squared test
res <- chisq.test(table_sex_group)
res # X-squared = 5.5335e-31, df = 1, p-value = 1
  table_sex_group <- table(df_nomed$sex, df_nomed$group)
  table_sex_group
  summary(df_nomed$sex[df_nomed$group=="0"])
  summary(df_nomed$sex[df_nomed$group=="1"])
  # Perform the chi-squared test
  res <- chisq.test(table_sex_group)
  res # X-squared = 0, df = 1, p-value = 1

# BMI
shapiro.test(df$bmi) #  NOT normally distributed with  p < 0.05
res<-wilcox.test(df$bmi[df$group=="1"], df$bmi[df$group=="0"])
res # W = 1203.5, p-value = 0.04171

# Medication
table_med_group <- table(df$medication_brain, df$group)
# Perform the chi-squared test
res <- chisq.test(table_med_group)
res # X-squared = 22.061, df = 1, p-value = 2.641e-06

# BDI
shapiro.test(df$bdi) #  NOT normally distributed with  p < 0.05
res<-wilcox.test(df$bdi[df$group=="1"], df$bdi[df$group=="0"])
res # X-squared = 22.061, df = 1, p-value = 2.641e-06

# STAI-T
shapiro.test(df$stai_t) #   normally distributed with  p > 0.05
res<-t.test(df$stai_t[df$group=="1"], df$stai_t[df$group=="0"])
res #  t = 3.662, df = 77.319, p-value = 0.0004559

# STAI-S
shapiro.test(df$stai_s) #   NOT normally distributed with  p > 0.05
res<-wilcox.test(df$stai_s[df$group=="1"], df$stai_s[df$group=="0"])
res 

cor.test(df$stai_t, df$stai_s) # r = 0.6728456, p-value = 6.962e-13
cor.test(df$bdi, df$stai_s) # r = 0.6740379, p-value = 6.125e-13
cor.test(df$stai_t, df$bdi) # r = 0.7689492 , p-value < 2.2e-16

# HR
shapiro.test(df$mean_RR_interval_hep) #   normally distributed with  p > 0.05
res<-t.test(df$mean_RR_interval_hep[df$group=="1"], df$mean_RR_interval_hep[df$group=="0"])
res # t = -3.2748, df = 85.58, p-value = 0.001527

# HRV
shapiro.test(df$rmssd_hep) #  NOT normally distributed with  p < 0.05
res<-wilcox.test(df$rmssd_hep[df$group=="1"], df$rmssd_hep[df$group=="0"])
res # W = 568, p-value = 0.0008739

# ECG amptlidue
# Ensure numeric conversion
df$ecg_amp_250full_norm <- as.numeric(df$ecg_amp_250full_norm)
df$ecg_amp_285_norm     <- as.numeric(df$ecg_amp_285_norm)
df$ecg_amp_310_norm     <- as.numeric(df$ecg_amp_310_norm)
df$ecg_amp_185_norm     <- as.numeric(df$ecg_amp_185_norm)

# Run normality tests
shapiro.test(df$ecg_amp_250full_norm) # normally distributed with  p > 0.05
shapiro.test(df$ecg_amp_285_norm)     # NOT normally distributed with  p < 0.05
shapiro.test(df$ecg_amp_310_norm)     # NOT normally distributed with  p < 0.05
shapiro.test(df$ecg_amp_185_norm)     # normally distributed with  p > 0.05

# Test difference as a list
test_results <- list()

test_results[["ecg_amplitude_250full (FND vs HC)"]] <- t.test( df$ecg_amp_250full_norm[df$group == "1"], df$ecg_amp_250full_norm[df$group == "0"])$p.value

test_results[["ecg_amplitude_285 (FND vs HC)"]] <- wilcox.test(df$ecg_amp_285_norm[df$group == "1"],df$ecg_amp_285_norm[df$group == "0"])$p.value

test_results[["ecg_amplitude_310 (FND vs HC)"]] <- wilcox.test( df$ecg_amp_310_norm[df$group == "1"], df$ecg_amp_310_norm[df$group == "0"])$p.value

test_results[["ecg_amplitude_185 (FND vs HC)"]] <- t.test( df$ecg_amp_185_norm[df$group == "1"], df$ecg_amp_185_norm[df$group == "0"])$p.value

# Convert to data frame
p_values_df <- data.frame(comparison = names(test_results),p_value= unlist(test_results))
# Adjust for multiple comparisons
p_values_df$p_adj_fdr <- p.adjust(p_values_df$p_value, method = "fdr")
# Print results
print(p_values_df) # NO differences between groups



#### scholarity
Data.scholarity <- read.xlsx("~/Documents/GitHub/interoception_NS/data/processed/12-2024/Demographics_Interoception.xlsx")

scholarity_labels <- c(
  "none",
  "compulsory school (primary school; 9 years)",
  "apprenticeship",
  "Matura (cantonal highschool)",
  "Higher technical collage (HF or FH)",
  "bachelor's degree",
  "master's degree",
  "PhD",
  "other")

# Replace numeric codes (0–9; 3 is missing) with labels
Data.scholarity$highest_education <- factor(
  Data.scholarity$highest_education,
  levels = c(0, 1, 2, 4, 5, 6, 7, 8, 9),
  labels = scholarity_labels)

Data.scholarity$group <- factor(
  Data.scholarity$group,
  levels = c(0, 1),
  labels = c("HC", "FND"))

table_scholarity <- table(Data.scholarity$highest_education, Data.scholarity$group)
table_scholarity

# save the table as excel
df_table <- as.data.frame.matrix(table_scholarity)
df_table <- cbind(Scholarity = rownames(df_table), df_table)
write_xlsx(df_table, "1.1_scholarity_table.xlsx")

# statistical test
fisher_test <- fisher.test(table_scholarity)
fisher_test

```

# 1.2 Effect Sizes

```{r}
#| warning: FALSE
#| error: FALSE
# calculation of cohens d for variable of interest
# Define the function
calculate_cohens_d <- function(data, variable, group_var, group1, group2) {
  # Subset the data for each group
  group1_data <- data[[variable]][data[[group_var]] == group1]
  group2_data <- data[[variable]][data[[group_var]] == group2]
  
  # Calculate means and standard deviations
  mean1 <- mean(group1_data, na.rm = TRUE)
  mean2 <- mean(group2_data, na.rm = TRUE)
  sd1 <- sd(group1_data, na.rm = TRUE)
  sd2 <- sd(group2_data, na.rm = TRUE)
  
  # Calculate sample sizes
  n1 <- sum(!is.na(group1_data))
  n2 <- sum(!is.na(group2_data))
  
  # Calculate pooled standard deviation
  sd_pooled <- sqrt(((n1 - 1) * sd1^2 + (n2 - 1) * sd2^2) / (n1 + n2 - 2))
  
  # Calculate Cohen's d
  cohens_d <- (mean1 - mean2) / sd_pooled
  
  # Return Cohen's d
  return(cohens_d)}

# Calculate Cohen's d for variables of differences
cohens_d_bmi <- calculate_cohens_d(df, "bmi", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_bmi) # cohens d of 0.501; FND has HIGHER values than HC

cohens_d_hr <- calculate_cohens_d(df, "mean_RR_interval_hep", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_hr) # cohens d of  -0.6906093; FND has LOWER values than HC

cohens_d_rmssd <- calculate_cohens_d(df, "rmssd_hep", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_rmssd) # cohens d of  -0.7769747; FND has LOWER values than HC

cohens_d_bdi <- calculate_cohens_d(df, "bdi", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_bdi) # cohens d of  1.354997; FND has HIGHER values than HC

cohens_d_stai_t <- calculate_cohens_d(df, "stai_t", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_stai_t)  # cohens d of  0.795104; FND has HIGHER values than HC

cohens_d_stai_s <- calculate_cohens_d(df, "stai_s", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_stai_s)  # cohens d of  1.03966; FND has HIGHER values than HC

cohens_d_anx_dep_SUM <- calculate_cohens_d(df, "anx_dep_SUM", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_anx_dep_SUM)  # cohens d of  1.108273; FND has HIGHER values than HC

cohens_d_sdq <- calculate_cohens_d(df, "sdq", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_sdq) # cohens d of  1.563886; FND has HIGHER values than HC

cohens_d_ctq_total <- calculate_cohens_d(df, "ctq_total", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_ctq_total) # cohens d of  0.6925085; FND has HIGHER values than HC

# for binary variable (medication intake) we calculate the ODDS RATIO
model <- glm(group ~ psychotropic_medication, data = df, family = binomial)
summary(model)
# Extract CI and Odds Ratio
confint_logit <- confint(model)  # CI on the log-odds scale
confint_or <- exp(confint_logit) # Exponentiate to get CI for OR
odds_ratio <- exp(coef(model)) # Calculate Odds Ratios
results <- cbind(odds_ratio, confint_or) # Combine OR and CI
results
                      
```

# 1.3 Summary Table

```{r}
#| warning: FALSE
#| error: FALSE
df_tbl <- df %>%
  mutate(sex = factor(sex, levels = c("1", "2"), labels = c("female", "male"))) %>%
  mutate(group = factor(group, levels = c("0", "1"), labels = c("HC", "FND")))  %>%
  mutate(medication_brain = factor(medication_brain, levels = c("0", "1"), labels = c("no", "yes, intake of these medication"))) %>%
  mutate(menopause = factor(menopause, levels = c("0", "1"), labels = c("no", "yes, in menopause")))

# Table 1 for Demographics
Data.t <- select(df_tbl, c("group", "sex", "age", "medication_brain", "bmi", "mean_RR_interval_hep", "rmssd_hep", "bdi", "stai_s", "stai_t"))

tbl1 <- Data.t %>%
  tbl_summary(
    by = group, # Grouping by 'group'
    missing = "no",
    statistic = list(
      age = "{mean} ({sd})", 
      sex = "{n} ({p})", 
      medication_brain ="{n} ({p})", 
      bmi = "{median} ({IQR})", 
      mean_RR_interval_hep = "{mean} ({sd})",
      rmssd_hep = "{median} ({IQR})",
      bdi = "{median} ({IQR})",
      stai_s = "{median} ({IQR})", 
      stai_t = "{mean} ({sd})" 
    ),
    digits = list(age = 1, sex = 1, medication_brain = 1, menopause = 1, bmi = 1, stai_s = 1, stai_t = 1, ctq_total = 1, sdq = 1, mean_RR_interval_hep = 1, rmssd_hep = 1),
    label = list(
      age = "Age (mean)",
      sex = "Sex",
      medication_brain = "Opioids, benzodiazeptine, psychotropic or antiepileptic medication",
      bmi = "Body Mass Index (median)",
      mean_RR_interval_hep = "Heartrate (mean)",
      rmssd_hep = "Heartrate Variability (median)",
      bdi = "Depression (median)",
      stai_s = "State Anxiety (median)",
      stai_t = "Trait Anxiety (mean)"    )
  ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Demographics Overview**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(
    test = list(
      age = "t.test",  # normally distrbuted variables
      bmi = "wilcox.test",  # non-noramlly distributed variables
      mean_RR_interval_hep = "t.test", 
      rmssd_hep = "wilcox.test",
      bdi = "wilcox.test",
      stai_s = "wilcox.test", 
      stai_t = "t.test", 
      sex = "chisq.test", # equally distributed dichotomous variables 
      medication_brain = "fisher.test"  # not equally dichotomous distributed 
    ) ) %>%  add_overall(last = FALSE)

# Print the table
tbl1

# save the table as excel
	tbl1_df <- tbl1 %>% as_tibble()
	write_xlsx(tbl1_df, "1.3_Table_demographics.xlsx")

```

# 1.4 Interoceptive Sensibility

```{r}
# MAIA
shapiro.test(df$maia_total) # normally distributed 
res<-t.test(df$maia_total[df$group=="1"], df$maia_total[df$group=="0"])
res # t = -3.279, df = 67.778, p-value = 0.001646
cohens_d_maia <- calculate_cohens_d(df, "maia_total", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_maia) # d = -0.724678


# IAS
shapiro.test(df$ias) # NOT normally distributed with  p < 0.05
res<-wilcox.test(df$ias[df$group=="1"], df$ias[df$group=="0"])
res # W = 681, p-value = 0.01951
cohens_d_ias <- calculate_cohens_d(df, "ias", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_ias) # d = -0.6697022

## VISUALIZATION
plotBoxplotGroups <- function(df, var, tit) { 
  plot <- ggplot(df, aes(x = group, y = var, fill = group)) + 
    geom_boxplot(show.legend = TRUE) + # Show legend for boxplot
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 2) + # Increase dot size
    labs(y = "Score") + 
    theme_classic() + 
    scale_fill_manual(values = c("#868686FF", "#BB4038")) + 
    scale_x_discrete(labels = c("HC", "FND")) +  
    guides(fill = "none") + # Remove the legend
    ggtitle(tit) +
    xlab("")
  return(plot)}


p1 <- plotBoxplotGroups(df, (df$maia_total),"MAIA")+ stat_compare_means(method = "t.test", paired = FALSE, label.x = 1.3, size = 5)
p2 <- plotBoxplotGroups(df, (df$ias),"IAS")+ stat_compare_means(method = "wilcox.test", paired = FALSE, label.x = 1.3, size = 5)
grid.arrange(p1, p2, ncol = 2)

	pdf("1.4_Plot_MAIA_IAS.pdf")
	grid.arrange(p1, p2, ncol = 2)
	dev.off()


```

## 1.4.1 MAIA subscores

```{r}
#| warning: FALSE
#| error: FALSE
  maia<-dplyr::select(df, pcode, group, maia_attreg, maia_note, maia_distr, maia_worry, maia_aware, maia_sfreg, maia_body, maia_trust)
  
  #Make long data
  maia_long <- gather(maia, key = "subscale", value = "score", -pcode, -group) # make data set long
  maia_long$subscale<-as.factor(maia_long$subscale)
 maia_long$subscale <- factor(maia_long$subscale, levels = rev(levels(maia_long$subscale)))
  
 # label the variables correctly
  maia_long$score<-as.numeric(maia_long$score)
  maia_long$group<-as.factor(maia_long$group)
  maia_long$group <- factor(maia_long$group, levels = c(0, 1), labels = c("HC", "FND"))
  
  # Calculate mean, median, CI, SE
  library(dplyr)
  alpha=0.05
  maia.summary <- maia_long %>%
    group_by(group,subscale)%>%
    dplyr::summarise(
      se = sd(score) / sqrt(length(score)),
      t=qt((1-alpha)/2 + .5, length(score)-1),   # tend to 1.96 if sample size is big enough
      CI=t*se, 
      score = mean(score), .groups = 'drop') # you can use here mean or median depending on data distribution
  
  maia.summary # according to this order name the lables in the following plot p
  
  # STATISTICS
  # Multiple t-tests using FDR correction for multiple comparisons (FDR or bonferroni)
  library(rstatix)
  stat.test <- maia_long %>%
    group_by(subscale) %>%
    t_test(score ~ group) %>%
    adjust_pvalue(method = "fdr") %>% # correct for multiple comparision (as they are subscales)
    add_significance()
  stat.test # subscales worrying, trusting, body and attention regulation are sig difference across groups
  
  # Filter for significant results (both * and **)
  significant_results <- stat.test %>%
    filter(p.adj.signif %in% c("*", "**")) %>%
    select(subscale, p.adj.signif)  # Keep both subscale and significance level
  
  # VISUALIZATION
  # Plot
  library(ggsignif)
  maia.summary <- maia.summary %>%
    left_join(significant_results, by = "subscale") %>%
    mutate(significance = ifelse(is.na(p.adj.signif), "", p.adj.signif))
  
  dodge <- position_dodge(width=0.9)
  maiaplot <- maia.summary %>%
  ggplot(aes(y = score, x = subscale, ymin = score - CI, ymax = score + CI, fill = group)) +
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.3) +
  labs(x = "MAIA Subscales", y = "Scores") +
  scale_fill_manual(values = c("#868686FF", "#BB4038")) +
  ggtitle("MAIA - Interoceptive Awareness") +
  theme_classic(base_size = 9) +
  scale_x_discrete(labels = c("Not-Worrying", "Trusting", "Self-Regulation", "Noticing", 
                              "Not-Distracting", "Body Listening", "Emotional Awareness", "Attention Regulation")) +
  guides(fill = guide_legend(title = "Groups"))
  
  maiaplot
  
	pdf("1.4.1_MAIAsubscales.pdf")
	maiaplot
	dev.off()
  
```

# 2) Clinical Supplements

## 2.1 Clinical Commorbidities

```{r}
#| warning: FALSE
#| error: FALSE
# Select the relevant columns and include the calculated duration
Data.t <- dplyr::select(df, c("group", "pain_medication", "psychotropic_medication", "sleep_medication", "other_medication", "ADS_ADD", "psychiatric_disorder", "hormonal_dysfunction", "other_diagnosis", "contraception", "menopause", "mc_regularity"))

Data.t <- Data.t  %>%
  mutate(group = factor(group, levels = c("0", "1"), labels = c("HC", "FND")))

tbl2 <- Data.t %>%
  tbl_summary(
    by = group,
    missing = "no",
    statistic = list(
      pain_medication = "{n} ({p}%)",
      psychotropic_medication = "{n} ({p}%)",
      sleep_medication = "{n} ({p}%)",
      other_medication = "{n} ({p}%)",
      ADS_ADD = "{n} ({p}%)",
      psychiatric_disorder = "{n} ({p}%)",
      hormonal_dysfunction = "{n} ({p}%)",
      other_diagnosis = "{n} ({p}%)",
      contraception = "{n} ({p}%)",
      menopause = "{n} ({p}%)",
      mc_regularity = "{n} ({p}%)"
    ),
    label = list(
      pain_medication = "Intake of Pain Medication [yes]",
      psychotropic_medication = "Intake of Psychotropic Medication [yes]",
      sleep_medication = "Intake of Sleep Medication [yes]",
      other_medication = "Intake of Other Medication [yes]",
      ADS_ADD = "Comorbidity of ADS or AD(H)D [yes]",
      psychiatric_disorder = "Comorbidity of Psychiatric Disorder [yes]",
      hormonal_dysfunction = "Comorbidity of Hormonal Dysfunction [yes]",
      other_diagnosis = "Comorbidity of other Diagnosis [yes]",
      contraception = "Intake of Hormonal Contraception [yes]",
      menopause = "in Menopause [yes]",
      mc_regularity = "Regular, natural cycle [yes]"
    )
  ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Clinical Characteristics**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(test = everything() ~ "fisher.test") %>%
  add_overall(last = FALSE)

tbl2
  
  # save table
	tbl2_df <- tbl2 %>% as_tibble()
	write_xlsx(tbl1_df, "2.1_Table_clinical_characteristics.xlsx")

Data.t <- dplyr::select(df_female, c("group", "contraception", "menopause", "mc_regularity"))

Data.t <- Data.t  %>%
  mutate(group = factor(group, levels = c("0", "1"), labels = c("HC", "FND")))

tbl2.2 <- Data.t %>%
  tbl_summary(
    by = group,
    missing = "no",
    statistic = list(
      contraception = "{n} ({p}%)",
      menopause = "{n} ({p}%)",
      mc_regularity = "{n} ({p}%)" ),
    label = list(
      contraception = "Intake of Hormonal Contraception [yes]",
      menopause = "in Menopause [yes]",
      mc_regularity = "Regular, natural cycle [yes]") ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Clinical Characteristics**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(test = everything() ~ "fisher.test") %>%
  add_overall(last = FALSE)

tbl2.2

  # save table
	tbl2_df <- tbl2.2 %>% as_tibble()
	write_xlsx(tbl1_df, "2.1_Table_clinical_characteristics_female.xlsx")

```

## 2.2 FND subtypes

```{r}
#| warning: FALSE
#| error: FALSE
# create variable of duation of months
library(gtsummary)
library(dplyr)

# Select relevant columns from df_tbl
Data.t <- select(df_FND, c("sex", "diagnosis", "fds", "motor", "weakness", "sensory", "pppd", "cognitive"))

# Create the summary table
tbl3 <- Data.t %>%
  tbl_summary(
    by = sex,
    missing = "no",
    statistic = list(
      fds = "{n} ({p})",
      motor = "{n} ({p})",
      weakness = "{n} ({p})",
      sensory = "{n} ({p})",
      pppd = "{n} ({p})",
      cognitive = "{n} ({p})"),
    digits = list(duration_months = 1,fds = 1,motor = 1,weakness = 1,sensory = 1,pppd = 1,cognitive = 1),
    label = list(
      sex = "Sex",
      fds = "Functional Dissociative Seizures [yes]",
      motor = "motor + symptoms [yes]",
      weakness = "motor - symptoms [yes]",
      sensory = "sensory symptoms [yes]",
      pppd = "dizziness (PPPD) [yes]",
      cognitive = "cognitive symptoms [yes]"
    )
  ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Clinical Characteristics**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_overall(last = FALSE)

# Print the summary table
tbl3

  # save table
	tbl2_df <- tbl3 %>% as_tibble()
	write_xlsx(tbl1_df, "2.2_Table_FNDsubtypes.xlsx")

summary(df_FND$duration_months)
sd(df_FND$duration_months) 

## additional table for diagnosis (instead of symptoms)
df_FND$diagnosis

# Define diagnosis labels
diagnosis_labels <- c(
  "44.5 Functional/Dissociative Seizures",
  "44.4 motor FND",                       
  "44.4 motor FND",                       
  "44.6 sensory FND",                     
  "44.7 mixed FND")

# Recode diagnosis
df_FND$diagnosis <- dplyr::case_when(df_FND$diagnosis == 1 ~ diagnosis_labels[1],
  df_FND$diagnosis %in% c(2, 3) ~ diagnosis_labels[2],
  df_FND$diagnosis == 4 ~ diagnosis_labels[4],
  df_FND$diagnosis == 5 ~ diagnosis_labels[5],
  TRUE ~ "Other")

# Convert to factors
df_FND$diagnosis <- factor(df_FND$diagnosis)
df_FND$sex <- factor(df_FND$sex, levels = c(1, 2), labels = c("Female", "Male"))

# Create contingency table
table_diagnosis <- table(df_FND$diagnosis, df_FND$sex)
print(table_diagnosis)

# Convert to data frame
df_table <- as.data.frame.matrix(table_diagnosis)
df_table <- cbind(Diagnosis = rownames(df_table), df_table)
writexl::write_xlsx(df_table, "2.2_diagnosis_FND.xlsx")

```

# 3) HBT: group differences

```{r}
#| warning: FALSE
#| error: FALSE
# non paramatric wilcox test for NOT normally distributed data 
shapiro.test(df$hbt_accuracy_mean) # NOT normally distributed
res<-wilcox.test(df$hbt_accuracy_mean[df$group=="1"], df$hbt_accuracy_mean[df$group=="0"])
res #  p-value = 0.06046
cohens_d_hbt_mean <- calculate_cohens_d(df, "hbt_accuracy_mean", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_hbt_mean) #-0.4293561

shapiro.test(df$hbt_change_accuracy) # NOT normally distributed
res<-wilcox.test(df$hbt_change_accuracy[df$group=="1"], df$hbt_change_accuracy[df$group=="0"])
res  # no sig difference between groups

shapiro.test(df$hbt_confidence_mean) # NOT normally distributed
res<-wilcox.test(df$hbt_confidence_mean[df$group=="1"], df$hbt_confidence_mean[df$group=="0"])
res # no sig difference between groups

shapiro.test(df$hbt_change_confidence) #  normally distributed
res<-t.test(df$hbt_change_confidence[df$group=="1"], df$hbt_change_confidence[df$group=="0"])
res  # no sig difference between groups

shapiro.test(df$hbt_intensity_mean) # NOT normally distributed
res<-wilcox.test(df$hbt_intensity_mean[df$group=="1"], df$hbt_intensity_mean[df$group=="0"])
res# no sig difference between groups

```

## 3.1 Summary Table

```{r}
#| warning: FALSE
#| error: FALSE
Data.t <- select(df_tbl, c("group", "hbt_accuracy_mean", "hbt_confidence_mean", "hbt_intensity_mean", "hbt_easiness_mean", "hbt_change_accuracy", "hbt_change_confidence", "maia_total", "traitPE_maiaHBT", "ias", "traitPE_iasHBT"))
library(gtsummary)
tbl_hbt <- Data.t %>%
  tbl_summary(
    by = group,
    missing = "no",
    statistic = list(
      hbt_accuracy_mean = "{median} ({IQR})",
      hbt_confidence_mean = "{median} ({IQR})",
      hbt_intensity_mean = "{median} ({IQR})",
      hbt_easiness_mean = "{median} ({IQR})",
      hbt_change_accuracy = "{median} ({IQR})",
      hbt_change_confidence = "{median} ({IQR})",
      maia_total = "{median} ({IQR})",
      traitPE_maiaHBT = "{median} ({IQR})",
      ias = "{median} ({IQR})",
      traitPE_iasHBT = "{median} ({IQR})"    ),
    digits = list(hbt_accuracy_mean = 1, hbt_confidence_mean = 1, hbt_intensity_mean = 1, hbt_easiness_mean = 1, 
                  hbt_change_accuracy = 1, hbt_change_confidence = 1, hep_focus = 1, hep_difficulty = 1, hep_intensity = 1, 
                  maia_total = 1, traitPE_maiaHBT = 1,ias = 1, traitPE_iasHBT = 1),
    label = list(hbt_accuracy_mean = "HBT accuracy",
                 hbt_confidence_mean= "HBT confidence",
                 hbt_intensity_mean= "HBT intensity",
                 hbt_easiness_mean= "HBT easiness",
                 hbt_change_accuracy= "HBT change in accuracy",
                 hbt_change_confidence = "HBT change in confidence",
                 maia_total = "Interoceptive Awareness (MAIA)",
                 traitPE_maiaHBT= "Trait-Prediction-Error (MAIA-HBT)",
                 ias= "Interoceptive Accuracy Self-Report (IAS)",
                 traitPE_iasHBT = "Trait-Prediction-Error (IAS-HBT)")
  ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Interoception Overview**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(test = everything() ~ "wilcox.test") %>% 
  add_overall(last = FALSE)

tbl_hbt

  # Save as png table
  gt_table <- as_gt(tbl_hbt) # convert to gt table
  output_path <- file.path("3.1_Table_HBT.png")
  gt::gtsave(gt_table, filename = output_path)

  # Save as xlsx table
  df_table <- as_tibble(tbl_hbt)
  output_path <- file.path("3.1_Table_HBT.xlsx")
  write_xlsx(df_table, path = output_path)
  
```

## 3.2 Change across Condition

```{r}
#| warning: FALSE
#| error: FALSE

hbt <- dplyr::select(df, pcode, group, hbt_accuracy_neutral, hbt_accuracy_arousal) 

# Make data long
library(tidyverse)
library(dplyr)
long_data <- gather(hbt, key = "HBT", value = "score", -pcode, -group) # make data set long 
long_data$HBT<-plyr::revalue(long_data$HBT, c("hbt_accuracy_neutral"="Neutral","hbt_accuracy_arousal" ="Arousal"))
long_data$HBT<-as.factor(long_data$HBT)
long_data$HBT <- factor(long_data$HBT, levels = rev(levels(long_data$HBT)))

# faceted, separate graphs per condition
library(ggplot2)
HBTplot <- ggplot(long_data, aes(x = HBT, y = score, group = pcode, color = group, alpha = 0.8)) +
  geom_line() + 
  geom_point() +
  facet_grid(. ~ group) + # Zum nach Gruppe unterteilen
  labs(title = "Change in HBT - Neutral vs. Arousal",
       x = "Heartbeat Tracking Task",
       y = "Accuracy Score",
       color = "Group") +
  theme_minimal() +   
  stat_smooth(aes(group = group, color = group), color = "black") + # Brauchts für die CI
  stat_summary(aes(group = group, color = group), geom = "point", fun = mean, shape = 17, size = 3.5, color = "black")+ # zum mean machen (die dickere linie, kannst du auch noch dicker machen)
  scale_alpha(guide = 'none') +
  scale_color_manual(values = c("black", "#E95248"), labels = c("HC", "FND"))
HBTplot

	pdf("3.2_Plot_HBTchange_acc.pdf")
	HBTplot
	dev.off()
	
# statistics:
res_FND<-wilcox.test(df$hbt_accuracy_neutral[df$group=="1"], df$hbt_accuracy_neutral[df$group=="1"], paired = TRUE)
res_FND #  no sig difference between condition in FND
res_HC<-wilcox.test(df$hbt_accuracy_arousal[df$group=="0"], df$hbt_accuracy_arousal[df$group=="0"], paired = TRUE)#  no sig difference between condition in HC
res_HC

  # Capture stat output as text
  res_FND_text <- capture.output(print(res_FND))
  res_HC_text <- capture.output(print(res_HC))
  results_text <- c("### Wilcoxon Test: FND Group ###", res_FND_text, 
                  "", "### Wilcoxon Test: HC Group ###", res_HC_text)
  writeLines(results_text, "3.2_stats_HBTchange_acc.txt")

# same thing for confidence
hbt_confidence <- dplyr::select(df, pcode, group, hbt_confidence_neutral, hbt_confidence_arousal) 
long_data_confidence <- gather(hbt_confidence, key = "HBT", value = "score", -pcode, -group) # make data set long 
long_data_confidence$HBT<-revalue(long_data_confidence$HBT, c("hbt_confidence_neutral"="Neutral","hbt_confidence_arousal" ="Arousal"))
long_data_confidence$HBT<-as.factor(long_data_confidence$HBT)
long_data_confidence$HBT <- factor(long_data_confidence$HBT, levels = rev(levels(long_data_confidence$HBT)))

HBTplot <- ggplot(long_data_confidence, aes(x = HBT, y = score, group = pcode, color = group, alpha = 0.8)) +
  geom_line() + 
  geom_point() +
  facet_grid(. ~ group) + 
  labs(title = "Change in HBT confidence - Neutral vs. Arousal",
       x = "Heartbeat Tracking Task",
       y = "confidence Score",
       color = "Group") +
  theme_minimal() +   
  stat_smooth(aes(group = group, color = group), color = "black") + # Brauchts für die CI
  stat_summary(aes(group = group, color = group), geom = "point", fun = mean, shape = 17, size = 3.5, color = "black")+ 
  scale_alpha(guide = 'none')+
  scale_color_manual(values = c("black", "#E95248"), labels = c("HC", "FND")) # Custom labels for 0 and 1

	pdf("3.2_Plot_HBTchange_conf.pdf")
	HBTplot
	dev.off()
	
# statistics:
res_FND<-wilcox.test(df$hbt_confidence_neutral[df$group=="1"], df$hbt_confidence_arousal[df$group=="1"], paired = TRUE)
res_FND #  no sig difference between condition in FND
res_HC<-wilcox.test(df$hbt_confidence_neutral[df$group=="0"], df$hbt_confidence_arousal[df$group=="0"], paired = TRUE)#  no sig difference between condition in HC
res_HC

  # Capture stat output as text
  res_FND_text <- capture.output(print(res_FND))
  res_HC_text <- capture.output(print(res_HC))
  results_text <- c("### Wilcoxon Test: FND Group ###", res_FND_text, 
                  "", "### Wilcoxon Test: HC Group ###", res_HC_text)
  writeLines(results_text, "3.2_stats_HBTchange_conf.txt")


```

# 4) HEP: debriefing questions (control analysis)

```{r}
#| warning: FALSE
#| error: FALSE
hist(df$hep_difficulty_heart)
shapiro.test(df$hep_difficulty_heart)  # NOT normally distributed with  p < 0.05
shapiro.test(df$hep_difficulty_heart[df$group == "1"]) # within patients hep difficulty heart is NOT normally distributed with  p < 0.05
shapiro.test(df$hep_difficulty_heart[df$group == "0"]) # within controls hep difficulty heart is NOT normally distributed with  p < 0.05
summary(df$hep_difficulty_heart)
summary(df_FND$hep_difficulty_heart)
summary(df_HC$hep_difficulty_heart)

hist(df$hep_difficulty_sound)
shapiro.test(df$hep_difficulty_sound)  # NOT normally distributed with  p < 0.05
shapiro.test(df$hep_difficulty_sound[df$group == "1"]) # within patients hep difficulty sound is NOT normally distributed with  p < 0.05
shapiro.test(df$hep_difficulty_sound[df$group == "0"]) # within controls hep difficulty sound is NOT normally distributed with  p < 0.05
summary(df$hep_difficulty_sound)
summary(df_FND$hep_difficulty_sound)
summary(df_HC$hep_difficulty_sound)

# Create an empty list to store p-values
test_results <- list()

# Function to compute Wilcoxon test and store p-values
store_p_value <- function(name, group1, group2, paired=FALSE) {
  test_results[[name]] <<- wilcox.test(group1, group2, paired=paired)$p.value}

# Between-group comparisons
store_p_value("hep_difficulty_heart (FND vs HC)", df$hep_difficulty_heart[df$group=="1"], df$hep_difficulty_heart[df$group=="0"])
store_p_value("hep_difficulty_sound (FND vs HC)", df$hep_difficulty_sound[df$group=="1"], df$hep_difficulty_sound[df$group=="0"])

# Within-group (paired) comparisons
store_p_value("hep_difficulty (FND: sound vs heart)", df_FND$hep_difficulty_sound, df_FND$hep_difficulty_heart, paired=TRUE)
store_p_value("hep_difficulty (HC: sound vs heart)", df_HC$hep_difficulty_sound, df_HC$hep_difficulty_heart, paired=TRUE)

# Between-group comparisons for focus
store_p_value("hep_focus_heart (FND vs HC)", df$hep_focus_heart[df$group=="1"], df$hep_focus_heart[df$group=="0"])
store_p_value("hep_focus_sound (FND vs HC)", df$hep_focus_sound[df$group=="1"], df$hep_focus_sound[df$group=="0"])

# Within-group (paired) comparisons for focus
store_p_value("hep_focus (FND: sound vs heart)", df_FND$hep_focus_sound, df_FND$hep_focus_heart, paired=TRUE)
store_p_value("hep_focus (HC: sound vs heart)", df_HC$hep_focus_sound, df_HC$hep_focus_heart, paired=TRUE)

# Between-group comparisons for intensity
store_p_value("hep_intensity_heart (FND vs HC)", df$hep_intensity_heart[df$group=="1"], df$hep_intensity_heart[df$group=="0"])
store_p_value("hep_intensity_sound (FND vs HC)", df$hep_intensity_sound[df$group=="1"], df$hep_intensity_sound[df$group=="0"])

# Within-group (paired) comparisons for intensity
store_p_value("hep_intensity (FND: sound vs heart)", df_FND$hep_intensity_sound, df_FND$hep_intensity_heart, paired=TRUE)
store_p_value("hep_intensity (HC: sound vs heart)", df_HC$hep_intensity_sound, df_HC$hep_intensity_heart, paired=TRUE)

# Convert list to dataframe
p_values_df <- data.frame(
  Comparison = names(test_results),
  p_value = unlist(test_results))

# Apply FDR correction
p_values_df$p_adj_fdr <- p.adjust(p_values_df$p_value, method = "fdr")

# Print results
print(p_values_df)

  # Save as a text file
  write.table(p_values_df, "4_debriefing.txt", sep = "\t", row.names = FALSE, quote = FALSE)

# Calculating cohens d for diferent self-reported difficutly
cohens_d_difficulty_heart <- calculate_cohens_d(df, "hep_difficulty_heart", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_difficulty_heart) 

```

## 4.1 Summary Table

```{r}
#| warning: FALSE
#| error: FALSE
### Summary Table for debriefing questionnaire ---------------------------------------------
df_tbl <- df %>%
  mutate(group = factor(group, levels = c("0", "1"), labels = c("HC", "FND")))

Data.t <- select(df_tbl, c("group","hep_difficulty_heart", "hep_difficulty_sound", "hep_focus_heart", "hep_focus_sound", "hep_intensity_heart", "hep_intensity_sound"))
library(gtsummary)
library(gtsummary)
library(dplyr)

# Create the table summary
tbl_hep <- Data.t %>%
  tbl_summary(
    by = group, # Grouping by 'group'
    missing = "no",
    statistic = list(
      hep_difficulty_heart = "{median} ({IQR})", 
      hep_difficulty_sound = "{median} ({IQR})", 
      hep_focus_heart ="{median} ({IQR})", 
      hep_focus_sound = "{median} ({IQR})", 
      hep_intensity_heart = "{median} ({IQR})", 
      hep_intensity_sound = "{median} ({IQR})" ),
    digits = list(hep_difficulty_heart = 1, hep_difficulty_sound = 1, hep_focus_heart = 1, hep_focus_sound = 1, hep_intensity_heart = 1, hep_intensity_sound = 1),
    label = list(
      hep_difficulty_heart = "Easiness to perceive Heartbeat",
      hep_difficulty_sound = "Easiness to perceive Sound",
      hep_focus_heart = "Ability to focus on Heartbeat",
      hep_focus_sound = "Ability to focus on Sound",
      hep_intensity_heart = "Intensity of perceived Heartbeat",
      hep_intensity_sound = "Intensity of perceived Sound") ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**Control Question**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(
    test = list(
      hep_difficulty_heart = "wilcox.test", 
      hep_difficulty_sound = "wilcox.test", 
      hep_focus_heart = "wilcox.test", 
      hep_focus_sound = "wilcox.test",
      hep_intensity_heart = "wilcox.test", 
      hep_intensity_sound = "wilcox.test"  )) %>% 
  add_overall(last = FALSE)

# Print the table
tbl_hep

  # save table
	tbl1_df <- tbl_hep %>% as_tibble()
	write_xlsx(tbl1_df, "4.1_Table_debriefing.xlsx")

```

# 5) HEP : explorative approach

We extracted the individual scores for the significant time intervals and electrodes / GMD / GFP obtained from the EEG analysis (done in Carool and EEGpal)

TRACE ANALYSIS:

-   F1 for 185-210ms when focusing on heart

-   F8 for 540-550s when focusing on sound (supplements)

-   P6 for 620-630ms when focusing on sound (supplements)

```{r}
#| warning: FALSE
#| error: FALSE
### Getting cohens d for the differences in traces across group
res<-wilcox.test(df$F1_HEART_185[df$group=="1"], df$F1_HEART_185[df$group=="0"])
res #    
cohens_d_F1 <- calculate_cohens_d(df, "F1_HEART_185", "group", 1, 0) # group 1=FND, group=0 HC
print(cohens_d_F1) #  -0.8010937
```

## 5.1 Testing for Covariates

demographic covariates: testing whether differences will stay when controlling for covariates

### 5.1.1 GLOBAL : interaction and group effects GFP and GMD

GLOBAL ANALYSIS

-   main effect of CONDITION:

    -   GMD at 510-524ms

-   interaction effect of CONDITIONxGROUP:

    -   GFP at 286-298ms

    -   GMD at 310-321ms

Please note: The topographical differences as main effect condition and interaction were identified using RAGU software (randomization statistics) using a Topographical ANOVA. This cannot be extracted as one value (unlike the GFP for example). We nonetheless extracted the Global Map Dissimilarity SChore (GMD) - which is not identical to the TANOVA results, but an approach to operationalize topographical differences with one value. Hence, the results should be interepreted with care.

```{r}
#| warning: FALSE
#| error: FALSE
# for the Interaction Effects we need the long data format
# Subset and rename EEG variables
df_subset <- df |>
 select( pcode, group, sex, age, medication_brain, anx_dep_SUM, #covariates to add
  rmssd_hep, mean_RR_interval_hep, ecg_amp_310_norm, ecg_amp_285_norm, # ecg for later
  GMD_HEART_310, GMD_SOUND_310, # interaction
  GFP_HEART_285, GFP_SOUND_285, # interaction
  GMD_HEART_510, GMD_SOUND_510 ) # main effect group

# Reshape to long format for GMD_310
df_gmd_long <- df_subset %>%
  pivot_longer(cols = c(GMD_HEART_310, GMD_SOUND_310),
    names_to = "condition",
    values_to = "EEG_Value") %>%
  mutate(CONDITION = if_else(grepl("HEART", condition), "HEART", "SOUND"),
    EEG_Type = "GMD_310" )

# Reshape to long format for GFP_285
df_gfp_long <- df_subset %>%
  pivot_longer(cols = c(GFP_HEART_285, GFP_SOUND_285),
    names_to = "condition",
    values_to = "EEG_Value" ) %>%
  mutate(CONDITION = if_else(grepl("HEART", condition), "HEART", "SOUND"),
    EEG_Type = "GFP_285")

# GMD_310 LINEAR MIXED EFFECTS MODEL without covariates (random effect for subject included to account for the rpeated measures)
model_gmd <- lmer(
  EEG_Value ~ group * CONDITION + (1 | pcode),
  data = df_gmd_long)
summary(model_gmd)

# GMD_310 model LINEAR MIXED EFFECTS MODEL with covariates as fixed effects included
model_gmd_cov <- lmer(
  EEG_Value ~ group * CONDITION + sex + age  + rmssd_hep + mean_RR_interval_hep + (1 | pcode),
  data = df_gmd_long)
summary(model_gmd_cov)

# GMD_310 model LINEAR MIXED EFFECTS MODEL with ECG trace as covariates as fixed effects included
model_gmd_cov <- lmer(
  EEG_Value ~ group * CONDITION +  ecg_amp_310_norm + (1 | pcode),
  data = df_gmd_long)
summary(model_gmd_cov)

# GFP_285 LINEAR MIXED EFFECTS MODEL without covariates (random effect for subject included to account for the repeated measures)
model_gfp <- lmer(
  EEG_Value ~ group * CONDITION + (1 | pcode),
  data = df_gfp_long)
summary(model_gfp) # interaction effect: p = 0.0302

# with linear mixed effect models the inclusion of covariates does not lead to differences in values, as they control for between group differences (and within group it does not change)

#  GFP_285 model LINEAR MIXED EFFECTS MODEL with ECG trace as covariates as fixed effects included
model_gfp_cov <- lmer(
  EEG_Value ~ group * CONDITION + ecg_amp_285_norm + (1 | pcode),
  data = df_gfp_long)
summary(model_gfp_cov) # effect of ECG amplitude


# FULL GFP_285 model LINEAR MIXED EFFECTS MODEL with ALL covariates as fixed effects included
model_gfp_cov <- lmer(
  EEG_Value ~ group * CONDITION +  sex + age +  rmssd_hep + mean_RR_interval_hep  + medication_brain + anx_dep_SUM + (1 | pcode),
  data = df_gfp_long)
summary(model_gfp_cov) ## interaction still sig (along with sex)


############ SUPPLEMENTS
# Reshape to long format for GMD_510
df_gmdmain_long <- df_subset %>%
  pivot_longer(cols = c(GMD_HEART_510, GMD_SOUND_510),
    names_to = "condition",
    values_to = "EEG_Value" ) %>%
  mutate(CONDITION = if_else(grepl("HEART", condition), "HEART", "SOUND"),
    EEG_Type = "GMD_510")

# GMD_510 LINEAR MIXED EFFECTS MODEL without covariates (random effect for subject included to account for the rpeated measures)
model_gmdmain <- lmer(
  EEG_Value ~ group * CONDITION + (1 | pcode),
  data = df_gmdmain_long)
summary(model_gmdmain)

# GMD_510 model LINEAR MIXED EFFECTS MODEL with covariates as fixed effects included
model_gmdmain_cov <- lmer(
  EEG_Value ~ group * CONDITION + sex + age + (1 | pcode),
  data = df_gmdmain_long)
summary(model_gmdmain_cov) # sex as a main effect

# GMD_510 model LINEAR MIXED EFFECTS MODEL with covariates as fixed effects included
model_gmdmain_cov <- lmer(
  EEG_Value ~ group * CONDITION + rmssd_hep + mean_RR_interval_hep + (1 | pcode),
  data = df_gmdmain_long)
summary(model_gmdmain_cov) # no interaction

# GMD_510 model LINEAR MIXED EFFECTS MODEL with covariates as fixed effects included
model_gmdmain_cov <- lmer(
  EEG_Value ~ group * CONDITION + anx_dep_SUM + medication_brain + (1 | pcode),
  data = df_gmdmain_long)
summary(model_gmdmain_cov) # only medication brain


# FULL GMD_510 model LINEAR MIXED EFFECTS MODEL with ALL (full model) covariates as fixed effects included
model_gmdmain_cov <- lmer(
  EEG_Value ~ group * CONDITION + sex + age + rmssd_hep + mean_RR_interval_hep + anx_dep_SUM + medication_brain + (1 | pcode),
  data = df_gmdmain_long)
summary(model_gmdmain_cov) ### sex and medication


```
### 5.1.2 TRACE: Group differences in F1 (and F8 and P6 for control condition)

```{r}
#| warning: FALSE
#| error: FALSE

# For trace F1 of focusing on the heart
lm <- lm(formula = F1_HEART_185 ~ group + sex + age,
          data = df) 
summary(lm) # GROUP P = 0.000374

lm <- lm(formula = F1_HEART_185 ~ group + rmssd_hep + mean_RR_interval_hep, data = df) 
summary(lm) # GROUP p = 0.000754

lm <- lm(formula = F1_HEART_185 ~ group + ecg_amp_185_norm,
          data = df) 
summary(lm) # GROUP, p = 0.000278

###  PSYCHIATRIC
lm <- lm(formula = F1_HEART_185 ~ group + medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # GROUP, p = 5.69e-05

#### FULL MODEL (including all covariates to see whether this way the effect will remain as well - however this is not reported in mansucript; model might be overfittedD)
lm <- lm(formula = F1_HEART_185 ~ group + sex + age + rmssd_hep + mean_RR_interval_hep+ medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # GROUP P = 0.000176

############ SUPPLEMENTS
# For trace F8 of focusing on the sound
lm <- lm(formula = F8_SOUND_540 ~ group + sex + age,
          data = df) 
summary(lm) # GROUP, p = 0.00029

### Cardiac
lm <- lm(formula = F8_SOUND_540 ~ group +rmssd_hep + mean_RR_interval_hep,
          data = df) 
summary(lm) # GROUP, p = 0.00108

### Psychiatric
lm <- lm(formula = F8_SOUND_540 ~ group + medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # GROUP, p = 0.00416

### FULL MODEL (including all covariates to see whether this way the effect will remain as well - however this is not reported in mansucript; model might be overfittedD)
lm <- lm(formula = F8_SOUND_540 ~ group + sex + age + rmssd_hep + mean_RR_interval_hep + medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # GROUP still, p = 0.0108

# For trace P6 of focusing on the sound
lm <- lm(formula = P6_SOUND_620 ~ group + sex + age,
          data = df) 
summary(lm) # GROUP, p = 0.00145

### Cardiac
lm <- lm(formula = P6_SOUND_620 ~ group + rmssd_hep + mean_RR_interval_hep,
          data = df) 
summary(lm) # GROUP, p = 0.00574

### Psychiatric
lm <- lm(formula = P6_SOUND_620 ~ group + medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # NOT SIG ANYMORE: GROUP p= 0.0748

### FULL MODEL (including all covariates to see whether this way the effect will remain as well - however this is not reported in mansucript; model might be overfittedD)
lm <- lm(formula = P6_SOUND_620 ~ group + sex + age + rmssd_hep + mean_RR_interval_hep + medication_brain + anx_dep_SUM,
          data = df) 
summary(lm) # NOT SIG ANYMORE; GROUP p= 0.1168

```


## 5.2 Predicting Symptom Severity with EEG markers activity: FND only interoceptive condition

```{r}
#| warning: FALSE
#| error: FALSE
# predicting symptom severity -----
df_sfmdrs <- df_FND %>%
  filter(!diagnosis %in% c(1, 4, 6, 7)) # those with diagnosis PNSES (44.5), sensory FND (44.6), PPPD or cognitive FND only are excluded 
GMD_309_SFMDRS <- cor.test(df_sfmdrs$GMD_HEART_310, df_sfmdrs$sfmdrs, method = "pearson")
GMD_309_SFMDRS 
GMD_309_CGI <- cor.test(df_FND$GMD_HEART_310, df_FND$cgi, method = "pearson") 
GMD_309_CGI
GMD_309_SDQ <- cor.test(df_FND$GMD_HEART_310, df_FND$sdq, method = "pearson") 
GMD_309_SDQ 
GMD_309_pos <- cor.test(df_FND$GMD_HEART_310, df_FND$sum_positive_signs, method = "pearson") 
GMD_309_pos 

# Extract p-values
p_values <- c(GMD_309_SFMDRS$p.value, GMD_309_CGI$p.value, 
              GMD_309_SDQ$p.value, GMD_309_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

GMD_510_SFMDRS <- cor.test(df_sfmdrs$GMD_HEART_510, df_sfmdrs$sfmdrs, method = "pearson") 
GMD_510_SFMDRS 
GMD_510_CGI <- cor.test(df_FND$GMD_HEART_510, df_FND$cgi, method = "pearson") 
GMD_510_CGI
GMD_510_SDQ <- cor.test(df_FND$GMD_HEART_510, df_FND$sdq, method = "pearson") 
GMD_510_SDQ # p = 0.000239
GMD_510_pos <- cor.test(df_FND$GMD_HEART_510, df_FND$sum_positive_signs, method = "pearson") 
GMD_510_pos 

# Extract p-values
p_values <- c(GMD_510_SFMDRS$p.value, GMD_510_CGI$p.value, 
              GMD_510_SDQ$p.value, GMD_510_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr # 510-SDQ still 0.00095


# SDQ 
lm <- lm(formula = sdq ~ GMD_HEART_510 + sex + age,data = df_FND) 
summary(lm) # GMD_HEART_510

## control for covariates of affective sympotms and medication intake
lm <- lm(formula = sdq ~ GMD_HEART_510 + anx_dep_SUM + medication_brain,
          data = df_FND) 
summary(lm) # GMD_HEART_510


# Visulization of GMD_HEART_510 and SDQ 
GMD_HEART_510_sdq<- ggplot(df, aes(x = sdq, y = GMD_HEART_510, color = as.factor(group))) +
  geom_point() + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(x = "Somatoform Dissociation Questionnaire (SDQ-20)", 
       y = "GMD at 510-524ms post R peak during Focusing on the Heart", 
       title = "EEG marker for cardiovascular interoception and dissociation scores",
       color = "Group") +
 theme_minimal() +
  theme(panel.grid = element_blank()) +  # Remove gridlines
  scale_color_manual(values = c("FND" = "#E95248", "HC" = "#868686FF"))
GMD_HEART_510_sdq

	pdf("5.2_Plot_GMD_heart510_sdq.pdf")
	GMD_HEART_510_sdq
	dev.off()

df$GFP_SOUND_285
GFP_HEART_285_SFMDRS <- cor.test(df_sfmdrs$GFP_HEART_285, df_sfmdrs$sfmdrs, method = "pearson") 
GFP_HEART_285_SFMDRS 
GFP_HEART_285_CGI <- cor.test(df_FND$GFP_HEART_285, df_FND$cgi, method = "pearson") 
GFP_HEART_285_CGI
GFP_HEART_285_SDQ <- cor.test(df_FND$GFP_HEART_285, df_FND$sdq, method = "pearson") 
GFP_HEART_285_SDQ 
GFP_HEART_285_pos <- cor.test(df_FND$GFP_HEART_285, df_FND$sum_positive_signs, method = "pearson") 
GFP_HEART_285_pos 

# Extract p-values
p_values <- c(GFP_HEART_285_SFMDRS$p.value, GFP_HEART_285_CGI$p.value, 
              GFP_HEART_285_SDQ$p.value, GFP_HEART_285_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

F1_HEART_185_SFMDRS <- cor.test(df_sfmdrs$F1_HEART_185, df_sfmdrs$sfmdrs, method = "pearson") 
F1_HEART_185_SFMDRS 
F1_HEART_185_CGI <- cor.test(df_FND$F1_HEART_185, df_FND$cgi, method = "pearson") 
F1_HEART_185_CGI
F1_HEART_185_SDQ <- cor.test(df_FND$F1_HEART_185, df_FND$sdq, method = "pearson") 
F1_HEART_185_SDQ 
F1_HEART_185_pos <- cor.test(df_FND$F1_HEART_185, df_FND$sum_positive_signs, method = "pearson") 
F1_HEART_185_pos 

# Extract p-values
p_values <- c(F1_HEART_185_SFMDRS$p.value, F1_HEART_185_CGI$p.value, 
              F1_HEART_185_SDQ$p.value, F1_HEART_185_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

```

Visualization GFP plot

```{r}
#| warning: FALSE
#| error: FALSE
# Reshape to long format
df_gfp_plot <- df %>%
  select(pcode, group, GFP_HEART_285, GFP_SOUND_285) %>%
  pivot_longer(cols = c(GFP_HEART_285, GFP_SOUND_285),
               names_to = "Condition",
               values_to = "Value")
df_gfp_plot <- df_gfp_plot %>%
  mutate(group = factor(group, levels = c(0, 1), labels = c("HC", "FND")))

# Calulate mean and se
summary_df <- df_gfp_plot %>%
  group_by(group, Condition) %>%
  dplyr::summarise(
    mean = mean(Value, na.rm = TRUE),
    se = sd(Value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop")

# group colors
group_colors <- c("FND" = "#E95248", "HC" = "#868686FF")

# Relabel conditions for x-axis
summary_df <- summary_df %>%
  mutate(Condition = case_when(
    Condition == "GFP_HEART_285" ~ "focusing on HEART",
    Condition == "GFP_SOUND_285" ~ "focusing on SOUND",
    TRUE ~ as.character(Condition)))

#  bar plot with SE error bars
ggplot(summary_df, aes(x = Condition, y = mean, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(0.9), width = 0.2) +
  scale_fill_manual(values = group_colors) +
  expand_limits(y = 0.5) +  # Soft lower limit, won't cut off bars
  labs(y = "Mean GFP", x = NULL, fill = "Group") +
  theme_minimal() +
  theme( panel.grid = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "right")

```

## 5.3 Correlation of interoceptive EEG markers and other interoceptive variables

```{r}
#| warning: FALSE
#| error: FALSE
# correlation  between neuronal markers and interoceptive scores
df_exclusion <- df[df$pcode != "63", ] # outlier in RRST

GMD_309_IAS <- cor.test(df$GMD_HEART_310, df$ias, method = "pearson") 
GMD_309_IAS 
GMD_309_MAIA <- cor.test(df$GMD_HEART_310, df$maia_total, method = "pearson") 
GMD_309_MAIA
GMD_309_rrst <- cor.test(df_exclusion$GMD_HEART_310, df_exclusion$rrst_sensitivity, method = "pearson") 
GMD_309_rrst
GMD_309_hbt<- cor.test(df$GMD_HEART_310, df$hbt_accuracy_mean, method = "pearson") 
GMD_309_hbt 

# Extract p-values
p_values <- c(GMD_309_IAS$p.value, GMD_309_MAIA$p.value, 
              GMD_309_rrst$p.value, GMD_309_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

GMD_510_IAS <- cor.test(df$GMD_HEART_510, df$ias, method = "pearson") 
GMD_510_IAS # 0.05882
GMD_510_MAIA <- cor.test(df$GMD_HEART_510, df$maia_total, method = "pearson") 
GMD_510_MAIA
GMD_510_rrst <- cor.test(df_exclusion$GMD_HEART_510, df_exclusion$rrst_sensitivity, method = "pearson") 
GMD_510_rrst # r = -0.28, p-value = 0.008867
GMD_510_hbt<- cor.test(df$GMD_HEART_510, df$hbt_accuracy_mean, method = "pearson") 
GMD_510_hbt # p-value = 0.01715; r = -0.25 GFP 289 and HBT accuracy mean

# Extract p-values
p_values <- c(GMD_510_IAS$p.value, GMD_510_MAIA$p.value, 
              GMD_510_rrst$p.value, GMD_510_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr # RRST stil p = 0.034, HBT still p = 0.034

# GMD_HEART_510_sdq
GMD_HEART_510_rrst<- ggplot(df_exclusion, aes(x = rrst_sensitivity, y = GMD_HEART_510, color = as.factor(group))) +
  geom_point() +  
  geom_smooth(method = "lm", se = TRUE) +  # Add regression lines and SE
  labs(x = "RRST sensitivity score", 
       y = "GMD at 510-524ms post R peak during Focusing on the Heart", 
       title = "EEG marker for cardiovascular interoception and other interoceptive scores",
       color = "Group") +  
 theme_minimal() +
  theme(panel.grid = element_blank()) +  # Remove gridlines
  scale_color_manual(values = c("1" = "#E95248", "0" = "#868686FF"))
GMD_HEART_510_rrst

	pdf("5.3_Plot_GMD_heart510_rrst.pdf")
	GMD_HEART_510_rrst
	dev.off()
	
# GMD_HEART_510_sdq
GMD_HEART_510_hbt<- ggplot(df, aes(x = hbt_accuracy_mean, y = GMD_HEART_510, color = as.factor(group))) +
  geom_point() +  
  geom_smooth(method = "lm", se = TRUE) +  
  labs(x = "Cardiovascular Interoceptive Accuracy (HBT)", 
       y = "GMD at 510-524ms post R peak during Focusing on the Heart", 
       title = "EEG marker for cardiovascular interoception and other interoceptive scores",
       color = "Group") +  # Label for the legend
 theme_minimal() +
  theme(panel.grid = element_blank()) +  # Remove gridlines
  scale_color_manual(values = c("FND" = "#E95248", "HC" = "#868686FF"))
GMD_HEART_510_hbt

	pdf("5.3_Plot_GMD_heart510_hbt.pdf")
	GMD_HEART_510_hbt
	dev.off()

GFP_285_IAS <- cor.test(df$GFP_HEART_285, df$ias, method = "pearson") 
GFP_285_IAS 
GFP_285_MAIA <- cor.test(df$GFP_HEART_285, df$maia_total, method = "pearson") 
GFP_285_MAIA
GFP_285_rrst <- cor.test(df_exclusion$GFP_HEART_285, df_exclusion$rrst_sensitivity, method = "pearson") 
GFP_285_rrst
GFP_285_hbt<- cor.test(df$GFP_HEART_285, df$hbt_accuracy_mean, method = "pearson") 
GFP_285_hbt

# Extract p-values
p_values <- c(GFP_285_IAS$p.value, GFP_285_MAIA$p.value, 
              GFP_285_rrst$p.value, GFP_285_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

F1_185_IAS <- cor.test(df$F1_HEART_185, df$ias, method = "pearson") 
F1_185_IAS #p-value = 0.04296, r = 0.2162987 
F1_185_MAIA <- cor.test(df$F1_HEART_185, df$maia_total, method = "pearson") 
F1_185_MAIA # p-value = 0.08494
F1_185_rrst <- cor.test(df$F1_HEART_185, df$rrst_sensitivity, method = "pearson") 
F1_185_rrst
F1_185_hbt<- cor.test(df$F1_HEART_185, df$hbt_accuracy_mean, method = "pearson") 
F1_185_hbt

# Extract p-values
p_values <- c(F1_185_IAS$p.value, F1_185_MAIA$p.value, 
              F1_185_rrst$p.value, F1_185_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr # not sig anymore


# visualization
F1_185_IAS<- ggplot(df, aes(x = ias, y = F1_HEART_185, color = as.factor(group))) +
  geom_point() +  
  geom_smooth(method = "lm", se = TRUE) + 
  labs(x = "Interoceptive Accuracy Scale (IAS)", 
       y = "F1 at 185-210ms post R peak during Focusing on the Heart", 
       title = "EEG marker for cardiovascular interoception and interoceptive sensibility",
       color = "Group") + 
 theme_minimal() +
  theme(panel.grid = element_blank()) +  # Remove gridlines
  scale_color_manual(values = c("FND" = "#E95248", "HC" = "#868686FF"))
F1_185_IAS

	pdf("5.3_Plot_F1_185_IAS.pdf")
	F1_185_IAS
	dev.off()
```

# 6) HEP : targeted approach

For the targeted approach, the three electrodes Fz, F8 and F7 were chosen based on previous FND literature for the time interval of 250-595ms after Rpeak. Trace differences were tested in Cartool. Here, we extracted the trace amplitudes in order to run interaction and correlation analysis.

## 6.1 Full time interval for F7, F8, Fz; 250-595ms

```{r}
#| warning: FALSE
#| error: FALSE
## LOAD  EEG DATA  -----------------------------------------------------------
# load prepared and cleaned file with averaged electrodes (separate per group and condition for the selected electrodes; F7, F8 and Fz but also the control electrodes)
df_EEG_FND <- read.xlsx("/Users/nataschastoffel/Documents/GitHub/interoception_NS/data/processed/02-2025/AverageTimeInterval_717-1423tf_FND_electrodes_250-595ms.xlsx")
df_EEG_HC <- read.xlsx("/Users/nataschastoffel/Documents/GitHub/interoception_NS/data/processed/02-2025/AverageTimeInterval_717-1423tf_HC_electrodes_250-595ms.xlsx")

df_EEG_FND$pcode <- substr(df_EEG_FND$pcode, 1, 4) # keeps only the first four digits (P001 etc) in the names
df_EEG_FND$pcode <- sub("^P00", "", df_EEG_FND$pcode) # gets rid of the P00 for all participants
df_EEG_FND$pcode <- sub("^P0", "", df_EEG_FND$pcode) # gets rid of the P0 for all participants
df_EEG_FND$pcode <- sub("^P", "", df_EEG_FND$pcode) # gets rid of the P for all participants

df_EEG_HC$pcode <- substr(df_EEG_HC$pcode, 1, 4) # keeps only the first four digits (P001 etc) in the names
df_EEG_HC$pcode <- sub("^P00", "", df_EEG_HC$pcode) # gets rid of the P00 for all participants
df_EEG_HC$pcode <- sub("^P0", "", df_EEG_HC$pcode) # gets rid of the P0 for all participants
df_EEG_HC$pcode <- sub("^P", "", df_EEG_HC$pcode) # gets rid of the P for all participants


# Add GROUP variable and merge datasets so we can afterwards do a long format
df_EEG_FND <- df_EEG_FND %>% mutate(GROUP = "FND")
df_EEG_HC <- df_EEG_HC %>% mutate(GROUP = "HC")

df_EEG<-merge(df_EEG_HC, df_EEG_FND, x.by=c("pcode"),y.by=c("pcode"),all.x=TRUE, all.y=TRUE) 

# Convert from wide to long format
df_long <- df_EEG %>%
  pivot_longer(cols = c(Fp1_HEART,	F3_HEART,	F7_HEART,	C3_HEART,	Fpz_HEART,	Fp2_HEART,	Fz_HEART,	F4_HEART,	F8_HEART,	Cz_HEART,	C4_HEART,	Fp1_SOUND,	F3_SOUND,	F7_SOUND,	C3_SOUND,	Fpz_SOUND,	Fp2_SOUND,	Fz_SOUND,	F4_SOUND,	F8_SOUND,	Cz_SOUND,	C4_SOUND),
               names_to = c("Electrode", "CONDITION"),
               names_pattern = "(F7|Fz|F8|Fp1|F3|C3|Fpz|Fp2|F4|Cz|C4)_(SOUND|HEART)",
               values_to = "EEG_Value") %>%
  mutate(GROUP = factor(GROUP),
         CONDITION = factor(CONDITION),
         Electrode = factor(Electrode),
         pcode = factor(pcode))  # Ensure participant is a factor

## merge with ECG data (to control for its effect in the ANOVA)
df_ECG <- df %>%
  select(pcode, sex, age, mean_RR_interval_hep, rmssd_hep, ecg_amp_250full_norm, medication_brain, anx_dep_SUM) %>%
  mutate(pcode = as.factor(pcode))  # Convert pcode to factor

df_long <- df_long %>%
  left_join(df_ECG %>% select(pcode, sex, age, rmssd_hep, mean_RR_interval_hep, ecg_amp_250full_norm, medication_brain, anx_dep_SUM), by = "pcode")


# Run repeated measures ANOVA for each electrode separately 
# aov for ANOVA with repeated measures, with the Error () term modeling within-subject variances

# Run ANOVA for F7
df_F7 <- df_long %>% filter(Electrode == "F7")
# FIRST Test normality by group and condition
df_F7  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # normality violated
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_F7) # homogenity assumption met
### SUMMARY of assumption test: CAREFUL in interpretation  as non-normal distribution


# now run ANOVA for F7
model_F7 <- aov(EEG_Value ~ GROUP * CONDITION + Error(pcode / CONDITION), data = df_F7)
summary(model_F7)
# no sig effect of condition or group*condition interaction

model_F7 <- aov(EEG_Value ~ GROUP * CONDITION + sex + age + Error(pcode / CONDITION), data = df_F7)
summary(model_F7)
# no sig effect of condition or group*condition interaction

df_F7 <- df_F7 %>% drop_na(rmssd_hep, mean_RR_interval_hep)
model_F7_ECG <- aov(EEG_Value ~ GROUP * CONDITION +  rmssd_hep + mean_RR_interval_hep + Error(pcode / CONDITION), data = df_F7)
summary(model_F7_ECG)
# no sig effect of group, HRV or HR; also no sig effect of condition or group*condition interaction
model_F7_ECG <- aov(EEG_Value ~ GROUP * CONDITION +  ecg_amp_250full_norm
                    + Error(pcode / CONDITION), data = df_F7)
summary(model_F7_ECG)


# Run ANOVA for Fz
df_Fz <- df_long %>% filter(Electrode == "Fz")
# FIRST Test normality by group and condition
df_Fz  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # FND sound condition not normally distributed
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_Fz) # homogenity assumption met
#SUMMARY of assumption test: only one condition not normally distributed (--> caution!) but proceeding with ANOVA

# Now, run ANOVA
model_Fz <- aov(EEG_Value ~ GROUP * CONDITION + Error(pcode / CONDITION), data = df_Fz)
summary(model_Fz)
# no sig effect of condition or group*condition interaction

model_Fz <- aov(EEG_Value ~ GROUP * CONDITION + sex + age + Error(pcode / CONDITION), data = df_Fz)
summary(model_Fz)

model_Fz_ECG <- aov(EEG_Value ~ GROUP * CONDITION + rmssd_hep + mean_RR_interval_hep + Error(pcode / CONDITION), data = df_Fz)
summary(model_Fz_ECG)
# no sig effect of group, HRV or HR (trend for HRV); also no sig effect of condition or group*condition interaction
model_Fz_ECG <- aov(EEG_Value ~ GROUP * CONDITION + ecg_amp_250full_norm
                    + Error(pcode / CONDITION), data = df_Fz)
summary(model_Fz_ECG)


# Run ANOVA for F8 
df_F8 <- df_long %>% filter(Electrode == "F8")
# FIRST Test normality by group and condition
df_F8  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value)# non-normality for FND SOUND and HC HEART (p < 0.05)
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_F8) # p = 0.945 indicates equal vairance across groups; the homogeneity assumption is satisfied
## SUMMARY of assumptions testing: Since ANOVA is robust to mild non-normality and groups N > 20 per group, and you have equal variances, we  proceed with ANOVA

# Now, run ANOVA
model_F8 <- aov(EEG_Value ~ GROUP * CONDITION + Error(pcode / CONDITION), data = df_F8)
summary(model_F8)
#  sig effect of condition but no sig group*condition interaction

model_F8_cov <- aov(EEG_Value ~ GROUP * CONDITION + sex + age + Error(pcode / CONDITION), data = df_F8)
summary(model_F8_cov) # group effect remains

model_F8_ecg <- aov(EEG_Value ~ GROUP * CONDITION + rmssd_hep + mean_RR_interval_hep + Error(pcode / CONDITION), data = df_F8)
summary(model_F8_ecg)

model_F8_psych <- aov(EEG_Value ~ GROUP * CONDITION + medication_brain + anx_dep_SUM + Error(pcode / CONDITION), data = df_F8)
summary(model_F8_psych)

model_F8_full<- aov(EEG_Value ~ GROUP * CONDITION + sex + age + rmssd_hep + mean_RR_interval_hep + medication_brain + anx_dep_SUM + Error(pcode / CONDITION), data = df_F8)
summary(model_F8_full) ### MAIN EFFECT group survives also in FULL model (including all variables in one model); p = 0.023

model_F8_ECG <- aov(EEG_Value ~ GROUP * CONDITION + ecg_amp_250full_norm  + Error(pcode / CONDITION), data = df_F8)
summary(model_F8_ECG)


###### running non-normality robust ARTool (instead of ANOVA)
# ART model
art_model_f8 <- ARTool::art(EEG_Value ~ GROUP * CONDITION + (1 | pcode), data = df_F8)
anova(art_model_f8) # main effect of group : F(1, 86) = 5.62, p = 0.020
### problem: not possible to include covariates, so only main group*condition effects are tested
```

## 6.2 Control Analysis: 2x3x2 analysis

```{r}
#| warning: FALSE
#| error: FALSE
# REGION 0 (our target electrodes; F7, Fz, F8)
df_region0 <- df_long %>% dplyr::filter(Electrode %in% c("F7", "Fz", "F8"))

# FIRST Test normality by group and condition
df_region0  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # p< 0.005; normality violated
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_region0) # homogenity assumption met (p>0.005)
### SUMMARY of assumption test: CAREFUL in interpretation  as it is non-normal distribution

# Now repeated-measures ANOVA (mixed design) with Error term: subject nested within the within-subject factors REGION and CONDITION
model_mixed <- aov(
  EEG_Value ~ GROUP * Electrode * CONDITION + Error(pcode / (Electrode * CONDITION)),
  data = df_region0)
summary(model_mixed)
#### significant main effect of GROUP and sig main effect of electrode (which is not suprirsing). However, there is not interaction of group and electrode, indicating similar electrode pattern across group, as well as similar condition effect accros group (no group X condition interaction)

###### running non-normality robust ARTool (instead of ANOVA)
# ART model
art_model_region0 <- ARTool::art(EEG_Value ~ GROUP * CONDITION * Electrode + Error(pcode / (Electrode * CONDITION)), data = df_region0)
anova(art_model_region0) 
## main effect of group and electorde remain (no interactions still)

# REGION 1 (Fp1, Fpz, Fp2)
df_region1 <- df_long %>% dplyr::filter(Electrode %in% c("Fp1", "Fz", "Fp2"))

# FIRST Test normality by group and condition
df_region1  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # p< 0.005; normality violated
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_region1) # homogenity assumption met (p>0.005)
### SUMMARY of assumption test: CAREFUL in interpretation  as it is non-normal distribution

# Now repeated-measures ANOVA (mixed design) with Error term: subject nested within the within-subject factors REGION and CONDITION
model_mixed <- aov(
  EEG_Value ~ GROUP * Electrode * CONDITION + Error(pcode / (Electrode * CONDITION)),
  data = df_region1)
summary(model_mixed)
#### significant main effect of electrode (non surprising; meaning EEG value varies across electrode site) but NOTHING (neither main nor interaction effect) with group

###### running non-normality robust ARTool (instead of ANOVA)
# ART model
art_model_region1 <- ARTool::art(EEG_Value ~ GROUP * CONDITION * Electrode + Error(pcode / (Electrode * CONDITION)), data = df_region1)
anova(art_model_region1) 
## still only main effect of electrode


# REGION 2 (F3, Fz, F4)
df_region2 <- df_long %>% dplyr::filter(Electrode %in% c("F3", "Fz", "F4"))

# FIRST Test normality by group and condition
df_region2  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # p< 0.005; normality violated
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_region2) # homogenity assumption met (p>0.005)
### SUMMARY of assumption test: CAREFUL in interpretation  as it is non-normal distribution

# Now repeated-measures ANOVA (mixed design) with Error term: subject nested within the within-subject factors REGION and CONDITION
model_mixed <- aov(
  EEG_Value ~ GROUP * Electrode * CONDITION + Error(pcode / (Electrode * CONDITION)),
  data = df_region2)
summary(model_mixed)
#### again only sig effect of electrode, and no interaction with neither group nor condition

###### running non-normality robust ARTool (instead of ANOVA)
# ART model
art_model_region2 <- ARTool::art(EEG_Value ~ GROUP * CONDITION * Electrode + Error(pcode / (Electrode * CONDITION)), data = df_region2)
anova(art_model_region2) 
## now we also have a sig main effect of GROUP, and still electrode, however no interaction still


# REGION 3 (C3, Cz, C4)
df_region3 <- df_long %>% dplyr::filter(Electrode %in% c("C3", "Cz", "C4"))

# FIRST Test normality by group and condition
df_region3  %>% group_by(GROUP, CONDITION) %>%  
  shapiro_test(EEG_Value) # p< 0.005; normality violated
# THEN, test homogenity of variance using Levene's test
car::leveneTest(EEG_Value ~ GROUP * CONDITION, data = df_region3) # homogenity assumption met (p>0.005)
### SUMMARY of assumption test: CAREFUL in interpretation  as it is non-normal distribution

# Now repeated-measures ANOVA (mixed design) with Error term: subject nested within the within-subject factors REGION and CONDITION
model_mixed <- aov(
  EEG_Value ~ GROUP * Electrode * CONDITION + Error(pcode / (Electrode * CONDITION)),
  data = df_region3)
summary(model_mixed)
#### Here we have a significant interaction between GROUP and CONDITION indicating that the EEG values differs between group, dependent on condition

### post-hoc test for this sig interaction:
library(emmeans)
emm <- emmeans(model_mixed, ~ GROUP * CONDITION)
pairs(emm, simple = "each")  # Only the FND group shows a significant condition effect: their EEG_Value is lower in the HEART condition than in SOUND (negative estimate). The HC group shows no change between conditions.
## Thus interaction driven by FND group differences

###### running non-normality robust ARTool (instead of ANOVA)
# ART model
art_model_region3 <- ARTool::art(EEG_Value ~ GROUP * CONDITION * Electrode + Error(pcode / (Electrode * CONDITION)), data = df_region3)
anova(art_model_region3) 
## only electrode as main effect, while conditionXgroup interaction now is only a trend p = 0.060
```

## 6.3 F8 separate per condition

Now, here we work with those time intervals that were sig in one condition only (as a specific group effect), testing whether the effect remains after control for demographics

-   536-547ms for HEART

-   570-581ms for HEART

-   451-462ms for SOUND

-   515-591ms for SOUND

```{r}
#| warning: FALSE
#| error: FALSE
## Listing pvalues of the identified group difference
# For trace F8 of focusing on the heart tiem interval 536-547
lm <- lm(formula = F8_HEART_536 ~ group,
          data = df) 
summary(lm) 

# For trace F8 of focusing on the heart tiem interval 570-581
lm <- lm(formula = F8_HEART_570 ~ group,
          data = df) 
summary(lm) 

# For trace F8 of focusing on the sound time interval 451-462
lm <- lm(formula = F8_SOUND_451 ~ group,
          data = df) 
summary(lm) 

# For trace F8 of focusing on the sound time interval 515-591
lm <- lm(formula = F8_SOUND_515 ~ group,
          data = df) 
summary(lm) 


# List to store p-values for the control analysis, and applying FDR correction
group_pvals <- c()
labels <- c()

# Define all formulas
formulas <- list(
  F8_HEART_536 ~ group + sex + age,
  F8_HEART_536 ~ group + rmssd_hep + mean_RR_interval_hep,
  F8_HEART_536 ~ group + medication_brain + anx_dep_SUM,
  
  F8_HEART_570 ~ group + sex + age,
  F8_HEART_570 ~ group + rmssd_hep + mean_RR_interval_hep,
  F8_HEART_570 ~ group + medication_brain + anx_dep_SUM,
  
  F8_SOUND_451 ~ group + sex + age,
  F8_SOUND_451 ~ group + rmssd_hep + mean_RR_interval_hep,
  F8_SOUND_451 ~ group + medication_brain + anx_dep_SUM,
  
  F8_SOUND_515 ~ group + sex + age,
  F8_SOUND_515 ~ group + rmssd_hep + mean_RR_interval_hep,
  F8_SOUND_515 ~ group + medication_brain + anx_dep_SUM)

# Run each model, extract group p-value
for (i in seq_along(formulas)) {
  model <- lm(formulas[[i]], data = df)
  smry <- summary(model)
  
  # Extract group coefficient p-value
  group_p <- coef(smry)["group1", "Pr(>|t|)"]
  group_pvals <- c(group_pvals, group_p)
  
  # Save label for identification
  labels <- c(labels, paste0("Model ", i, ": ", deparse(formulas[[i]])))}

# Apply FDR correction
fdr_pvals <- p.adjust(group_pvals, method = "fdr")

# Print results
for (i in seq_along(group_pvals)) {
  cat(labels[i], "\n")
  cat("  Uncorrected p:", sprintf("%.4f", group_pvals[i]),
      "| FDR-corrected p:", sprintf("%.4f", fdr_pvals[i]), "\n\n")}

```

## 6.4 Predicting Symptom Severity with EEG markers activity: FND only

```{r}
#| warning: FALSE
#| error: FALSE

### HEART 537ms
F8_HEART_536_SFMDRS <- cor.test(df_sfmdrs$F8_HEART_536, df_sfmdrs$sfmdrs, method = "pearson")
F8_HEART_536_SFMDRS 
F8_HEART_536_CGI <- cor.test(df_FND$F8_HEART_536, df_FND$cgi, method = "pearson") 
F8_HEART_536_CGI
F8_HEART_536_SDQ <- cor.test(df_FND$F8_HEART_536, df_FND$sdq, method = "pearson") 
F8_HEART_536_SDQ 
F8_HEART_536_pos <- cor.test(df_FND$F8_HEART_536, df_FND$sum_positive_signs, method = "pearson") 
F8_HEART_536_pos 

# Extract p-values
p_values <- c(F8_HEART_536_SFMDRS$p.value, F8_HEART_536_CGI$p.value, 
              F8_HEART_536_SDQ$p.value, F8_HEART_536_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

### HEART 570ms
F8_HEART_570_SFMDRS <- cor.test(df_sfmdrs$F8_HEART_570, df_sfmdrs$sfmdrs, method = "pearson")
F8_HEART_570_SFMDRS 
F8_HEART_570_CGI <- cor.test(df_FND$F8_HEART_570, df_FND$cgi, method = "pearson") 
F8_HEART_570_CGI
F8_HEART_570_SDQ <- cor.test(df_FND$F8_HEART_570, df_FND$sdq, method = "pearson") 
F8_HEART_570_SDQ 
F8_HEART_570_pos <- cor.test(df_FND$F8_HEART_570, df_FND$sum_positive_signs, method = "pearson") 
F8_HEART_570_pos 

# Extract p-values
p_values <- c(F8_HEART_570_SFMDRS$p.value, F8_HEART_570_CGI$p.value, 
              F8_HEART_570_SDQ$p.value, F8_HEART_570_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr


### F8_SOUND_451
F8_SOUND_451_SFMDRS <- cor.test(df_sfmdrs$F8_SOUND_451, df_sfmdrs$sfmdrs, method = "pearson")
F8_SOUND_451_SFMDRS 
F8_SOUND_451_CGI <- cor.test(df_FND$F8_SOUND_451, df_FND$cgi, method = "pearson") 
F8_SOUND_451_CGI
F8_SOUND_451_SDQ <- cor.test(df_FND$F8_SOUND_451, df_FND$sdq, method = "pearson") 
F8_SOUND_451_SDQ 
F8_SOUND_451_pos <- cor.test(df_FND$F8_SOUND_451, df_FND$sum_positive_signs, method = "pearson") 
F8_SOUND_451_pos 

# Extract p-values
p_values <- c(F8_SOUND_451_SFMDRS$p.value, F8_SOUND_451_CGI$p.value, 
              F8_SOUND_451_SDQ$p.value, F8_SOUND_451_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr


###  F8_SOUND_515
F8_SOUND_515_SFMDRS <- cor.test(df_sfmdrs$F8_SOUND_515, df_sfmdrs$sfmdrs, method = "pearson")
F8_SOUND_515_SFMDRS 
F8_SOUND_515_CGI <- cor.test(df_FND$F8_SOUND_515, df_FND$cgi, method = "pearson") 
F8_SOUND_515_CGI
F8_SOUND_515_SDQ <- cor.test(df_FND$F8_SOUND_515, df_FND$sdq, method = "pearson") 
F8_SOUND_515_SDQ 
F8_SOUND_515_pos <- cor.test(df_FND$F8_SOUND_515, df_FND$sum_positive_signs, method = "pearson") 
F8_SOUND_515_pos 

# Extract p-values
p_values <- c(F8_SOUND_515_SFMDRS$p.value, F8_SOUND_515_CGI$p.value, 
              F8_SOUND_515_SDQ$p.value, F8_SOUND_515_pos$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

```

## 6.5 Correlation of F8 and other Interoceptive measures

```{r}
#| warning: FALSE
#| error: FALSE
# Step 1: Reshape df_f8 from long to wide
df_exclusion <- df[df$pcode != "63", ] # outlier in RRST

### HEART 536ms
F8_HEART_536_ias <- cor.test(df$F8_HEART_536, df$ias, method = "pearson")
F8_HEART_536_ias 
F8_HEART_536_maia <- cor.test(df$F8_HEART_536, df$maia_total, method = "pearson") 
F8_HEART_536_maia
F8_HEART_536_rrst <- cor.test(df_exclusion$F8_HEART_536, df_exclusion$rrst_sensitivity, method = "pearson") 
F8_HEART_536_rrst 
F8_HEART_536_hbt <- cor.test(df$F8_HEART_536, df$hbt_accuracy_mean, method = "pearson") 
F8_HEART_536_hbt

# Extract p-values
p_values <- c(F8_HEART_536_ias$p.value, F8_HEART_536_maia$p.value, 
              F8_HEART_536_rrst$p.value, F8_HEART_536_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

### HEART 570ms
F8_HEART_570_ias <- cor.test(df$F8_HEART_570, df$ias, method = "pearson")
F8_HEART_570_ias 
F8_HEART_570_maia <- cor.test(df$F8_HEART_570, df$maia_total, method = "pearson") 
F8_HEART_570_maia
F8_HEART_570_rrst <- cor.test(df_exclusion$F8_HEART_570, df_exclusion$rrst_sensitivity, method = "pearson") 
F8_HEART_570_rrst 
F8_HEART_570_hbt <- cor.test(df$F8_HEART_570, df$hbt_accuracy_mean, method = "pearson") 
F8_HEART_570_hbt

# Extract p-values
p_values <- c(F8_HEART_570_ias$p.value, F8_HEART_570_maia$p.value, 
              F8_HEART_570_rrst$p.value, F8_HEART_570_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")

### F8_SOUND_451
F8_SOUND_451_ias <- cor.test(df$F8_SOUND_451, df$ias, method = "pearson")
F8_SOUND_451_ias 
F8_SOUND_451_maia <- cor.test(df$F8_SOUND_451, df$maia_total, method = "pearson") 
F8_SOUND_451_maia
F8_SOUND_451_rrst <- cor.test(df_exclusion$F8_SOUND_451, df_exclusion$rrst_sensitivity, method = "pearson") 
F8_SOUND_451_rrst 
F8_SOUND_451_hbt <- cor.test(df$F8_SOUND_451, df$hbt_accuracy_mean, method = "pearson") 
F8_SOUND_451_hbt

# Extract p-values
p_values <- c(F8_SOUND_451_ias$p.value, F8_SOUND_451_maia$p.value, 
              F8_SOUND_451_rrst$p.value, F8_SOUND_451_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr


###  F8_SOUND_515
F8_SOUND_515_ias <- cor.test(df$F8_SOUND_515, df$ias, method = "pearson")
F8_SOUND_515_ias 
F8_SOUND_515_maia <- cor.test(df$F8_SOUND_515, df$maia_total, method = "pearson") 
F8_SOUND_515_maia
F8_SOUND_515_rrst <- cor.test(df_exclusion$F8_SOUND_515, df_exclusion$rrst_sensitivity, method = "pearson") 
F8_SOUND_515_rrst 
F8_SOUND_515_hbt <- cor.test(df$F8_SOUND_515, df$hbt_accuracy_mean, method = "pearson") 
F8_SOUND_515_hbt

# Extract p-values
p_values <- c(F8_SOUND_515_ias$p.value, F8_SOUND_515_maia$p.value, 
              F8_SOUND_515_rrst$p.value, F8_SOUND_515_hbt$p.value)
# Apply FDR correction
p_values_fdr <- p.adjust(p_values, method = "fdr")
# Display corrected p-values
p_values_fdr

```

# 7) Control analysis: ECG

```{r}
#| warning: FALSE
#| error: FALSE
shapiro.test(df$mean_RR_interval_hep) # normally distributed with  p > 0.05
shapiro.test(df$rmssd_hep) #  NOT normally distributed with  p < 0.05

t.test(df$mean_RR_interval_hep[df$group=="1"], df$mean_RR_interval_hep[df$group=="0"])
t.test(df$rmssd_hep[df$group=="1"], df$rmssd_hep[df$group=="0"])

cohens_d_RR <- calculate_cohens_d(df, "mean_RR_interval_hep", "group", 1, 0) # group 1=FND, group=0 HC
cohens_d_RR
cohens_d_RMSSD <- calculate_cohens_d(df, "rmssd_hep", "group", 1, 0) # group 1=FND, group=0 HC
cohens_d_RMSSD


## LOAD  more specific ECG DATA  -----------------------------------------------------------
df_ECG <- read.xlsx("/Users/nataschastoffel/Documents/GitHub/interoception_NS/data/processed/02-2025/ECG_control_analysis.xlsx")

df_ECG$pcode <- substr(df_ECG$pcode, 1, 4) # keeps only the first four digits (P001 etc) in the names
df_ECG$pcode <- sub("^P00", "", df_ECG$pcode) # gets rid of the P00 for all participants
df_ECG$pcode <- sub("^P0", "", df_ECG$pcode) # gets rid of the P0 for all participants
df_ECG$pcode <- sub("^P", "", df_ECG$pcode) # gets rid of the P for all participants

# Ensure 'group' is a factor
df_ECG <- df_ECG %>%
  mutate(group = as.factor(group))  # Convert 'group' to factor

### SPECIFIC STATS per condition and group
# Create an empty list to store results
test_results <- list()

# Perform t-tests and Wilcoxon tests, storing p-values
test_results[["Mean_HEART (FND vs HC)"]] <- t.test(df_ECG$Mean_HEART[df_ECG$group=="FND"], df_ECG$Mean_HEART[df_ECG$group=="HC"])$p.value
test_results[["Mean_SOUND (FND vs HC)"]] <- t.test(df_ECG$Mean_SOUND[df_ECG$group=="FND"], df_ECG$Mean_SOUND[df_ECG$group=="HC"])$p.value

test_results[["RMSSD_HEART (FND vs HC)"]] <- wilcox.test(df_ECG$RMSSD_HEART[df_ECG$group=="FND"], df_ECG$RMSSD_HEART[df_ECG$group=="HC"])$p.value
test_results[["RMSSD_SOUND (FND vs HC)"]] <- wilcox.test(df_ECG$RMSSD_SOUND[df_ECG$group=="FND"], df_ECG$RMSSD_SOUND[df_ECG$group=="HC"])$p.value

test_results[["Mean_HEART vs Mean_SOUND (FND)"]] <- t.test(df_ECG$Mean_HEART[df_ECG$group=="FND"], df_ECG$Mean_SOUND[df_ECG$group=="FND"], paired=TRUE)$p.value
test_results[["Mean_HEART vs Mean_SOUND (HC)"]] <- t.test(df_ECG$Mean_HEART[df_ECG$group=="HC"], df_ECG$Mean_SOUND[df_ECG$group=="HC"], paired=TRUE)$p.value

test_results[["RMSSD_HEART vs RMSSD_SOUND (FND)"]] <- wilcox.test(df_ECG$RMSSD_HEART[df_ECG$group=="FND"], df_ECG$RMSSD_SOUND[df_ECG$group=="FND"], paired=TRUE)$p.value
test_results[["RMSSD_HEART vs RMSSD_SOUND (HC)"]] <- wilcox.test(df_ECG$RMSSD_HEART[df_ECG$group=="HC"], df_ECG$RMSSD_SOUND[df_ECG$group=="HC"], paired=TRUE)$p.value

# Convert list to dataframe
p_values_df <- data.frame(
  Comparison = names(test_results),
  p_value = unlist(test_results))

# Apply multiple comparison correction for heartrate and heartrate variability
p_values_df$p_adj_fdr <- p.adjust(p_values_df$p_value, method = "fdr")

# Print the results
print(p_values_df)



##### Testing interaction using reapeated measure ANOVA
# Convert Factor of 'Condition' to long format
df_long <- df_ECG %>%
  pivot_longer(cols = c(Mean_HEART, Mean_SOUND, RMSSD_HEART, RMSSD_SOUND),
               names_to = c(".value", "condition"), 
               names_pattern = "(.*)_(HEART|SOUND)") %>%
 mutate(
    condition = factor(condition, levels = c("HEART", "SOUND")),  # Ensure condition is a factor
    group = factor(group),  # Ensure group is a factor
    pcode = factor(pcode)   # Ensure participant ID is a factor
)

# ANOVA for Heartbeat
## FIRST check normality
df_long %>% group_by(group, condition) %>%
  shapiro_test(Mean) ## normality assumption met (p > 0.05)
## THEN check for homogenity of variance
car::leveneTest(Mean ~ group * condition, data = df_long) ## homogenity assumption met (p > 0.05)
## NOW run ANOVA for mean
model_Mean <- aov(Mean ~ group * condition + Error(pcode / condition), data = df_long)
summary(model_Mean)

# ANOVA for RMSSD
## FIRST check normality
df_long %>% group_by(group, condition) %>%
  shapiro_test(RMSSD) ## normality assumption NOT met for FND sound and both HC condition (p < 0.05)
## THEN check for homogenity of variance
car::leveneTest(RMSSD ~ group * condition, data = df_long) ## homogenity assumption NOT met (p = 0.0123)
## NOW run ANOVA for RMSSD
model_RMSSD <- aov(RMSSD ~ group * condition + Error(pcode / condition), data = df_long)
summary(model_RMSSD) ## group effect p = 0.0007

##### RUNNING a non-normality/homogenity robust analysis of ANOVA (ARTool) for RMSSD
library(ARTool)
# Fitting the ART model
art_model_rmssd <- art(RMSSD ~ group * condition + (1 | pcode), data = df_long)

# Run ANOVA on the aligned ranks
anova(art_model_rmssd) ### group effect p = 0.0014 

mean(df_ECG$Mean_HEART[df_ECG$group=="FND"]) # 836.9772ms 
mean(df_ECG$Mean_SOUND[df_ECG$group=="FND"]) # 832.7884ms --> faster heartbeat for SOUND

```

## 7.1 Summary Table

```{r}
#| warning: FALSE
#| error: FALSE
### Summary Table for debriefing questionnaire ---------------------------------------------
df_tbl <- df %>%
  mutate(group = factor(group, levels = c("0", "1"), labels = c("HC", "FND")))

Data.t <- select(df_tbl, c("group","mean_RR_interval_hep", "sd_RR_interval_hep", "rmssd_hep"))
library(gtsummary)
library(gtsummary)
library(dplyr)

# Create the table summary
tbl1 <- Data.t %>%
  tbl_summary(
    by = group, # Grouping by 'group'
    missing = "no",
    statistic = list(
      mean_RR_interval_hep = "{mean} ({sd})", 
      sd_RR_interval_hep = "{median} ({IQR})", 
      rmssd_hep ="{median} ({IQR})"
    ),
    
    digits = list(mean_RR_interval_hep = 1, sd_RR_interval_hep = 1, rmssd_hep = 1),
    label = list(
      mean_RR_interval_hep = "Mean RR interval in ms, mean (SD)",
      sd_RR_interval_hep = "SD of RR interval in ms, median(IQR)",
      rmssd_hep = "Heartrate Variability (RMSSD) in ms, median(IQR)"
      )
  ) %>%
  modify_header(list(label ~ "**Variable**")) %>%
  modify_caption("**ECG across group**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Group**") %>%
  add_p(
    test = list(
      mean_RR_interval_hep = "t.test", 
      sd_RR_interval_hep = "wilcox.test", 
      rmssd_hep = "wilcox.test"
    )
  ) %>% 
  add_overall(last = FALSE)

# Print the table
tbl1
	
  tbl1_df <- tbl1 %>% as_tibble()
	write_xlsx(tbl1_df, "7.1_Table_ECG.xlsx")
	
	

```

## 7.2 Correlations with EEG

```{r}
#| warning: FALSE
#| error: FALSE

# control ECG differences are assocaited with EEG markers

cor.test(df$mean_RR_interval_hep, df$GMD_HEART_310) # interaction explorative, global analysis
cor.test(df$mean_RR_interval_hep, df$GMD_HEART_510) # main effect group explorative, global analysis
cor.test(df$mean_RR_interval_hep, df$GFP_HEART_285) # interaction explorative, global analysis
cor.test(df$mean_RR_interval_hep, df$F1_HEART_185) # group explorative, trace analysis
cor.test(df$mean_RR_interval_hep, df$F8_SOUND_515) # group targeted, trace analysis
cor.test(df$mean_RR_interval_hep, df$F8_SOUND_451) # group targeted, trace analysis
cor.test(df$mean_RR_interval_hep, df$F8_HEART_536) # group targeted, trace analysis
cor.test(df$mean_RR_interval_hep, df$F8_HEART_570) # group targeted, trace analysis

cor.test(df$rmssd_hep, df$GMD_HEART_310) 
cor.test(df$rmssd_hep, df$GMD_HEART_510)
cor.test(df$rmssd_hep, df$GFP_HEART_285)
cor.test(df$rmssd_hep, df$F1_HEART_185)
cor.test(df$rmssd_hep, df$F8_SOUND_515)
cor.test(df$rmssd_hep, df$F8_SOUND_451)
cor.test(df$rmssd_hep, df$F8_HEART_536)
cor.test(df$rmssd_hep, df$F8_HEART_570)

# and also difficulty rating (that was different between groups)
cor.test(df$hep_difficulty_heart, df$GMD_HEART_310) # p=0.08
cor.test(df$hep_difficulty_heart, df$GMD_HEART_510)
cor.test(df$hep_difficulty_heart, df$GFP_HEART_285)
cor.test(df$hep_difficulty_heart, df$F1_HEART_185)
cor.test(df$hep_difficulty_heart, df$F8_SOUND_515)
cor.test(df$hep_difficulty_heart, df$F8_SOUND_451)
cor.test(df$hep_difficulty_heart, df$F8_HEART_536)
cor.test(df$hep_difficulty_heart, df$F8_HEART_570)
```

## 7.3 Visualization, specific for condition

```{r}
#| warning: FALSE
#| error: FALSE
# Reshape data to long format
df_long <- df_ECG %>%
  select(pcode, group, Mean_HEART, RMSSD_HEART, Mean_SOUND, RMSSD_SOUND) %>%
  pivot_longer(cols = -c(pcode, group), names_to = "Metric", values_to = "Value")

# Filter for Mean and RMSSD separately
df_mean <- df_long %>% filter(grepl("^Mean_", Metric))
df_rmssd <- df_long %>% filter(grepl("^RMSSD_", Metric))

# Mean Heart Rate Plot
p1 <- ggplot(df_mean, aes(x = Metric, y = Value, fill = group)) +
  geom_bar(stat = "summary", fun = "mean", position = position_dodge(0.8), alpha = 0.8, , width = 0.8) +
  geom_errorbar(stat = "summary", fun.data = mean_se, position = position_dodge(0.8), width = 0.2) +
  labs(title = "Heart Rate", y = "R-R interval in ms", x = "") +
  scale_fill_manual(values = c("FND" = "#E95248", "HC" = "#868686FF")) +
  coord_cartesian(ylim = c(600, NA)) +  # Zooms in without removing bars
   theme_minimal() +
  theme(legend.position = "none") +
  theme(panel.grid = element_blank())  # Remove grid lines

# RMSSD Plot
p2 <- ggplot(df_rmssd, aes(x = Metric, y = Value, fill = group)) +
  geom_bar(stat = "summary", fun = "mean", position = position_dodge(0.8), alpha = 0.8, , width = 0.8) + geom_errorbar(stat = "summary", fun.data = mean_se, position = position_dodge(0.8), width = 0.2) +
  labs(title = "Heart Rate Variability", y = "RMSSD in ms", x = "") +
  scale_fill_manual(values = c("FND" = "#E95248", "HC" = "#868686FF")) +
  theme_minimal() +
  theme(panel.grid = element_blank())  # Remove grid lines


# Print plots
grid.arrange(p1, p2, ncol = 2)

	pdf("7.3_Plot_ECG.pdf")
	grid.arrange(p1, p2, ncol = 2)
	dev.off()
	
```

## 7.4 Statistics, specific for condition

```{r}
#| warning: FALSE
#| error: FALSE
#### statistics
		# Paired comparisons: Reshape to wide format
df_wide <- df_ECG %>%
  select(pcode, group, Mean_HEART, Mean_SOUND, RMSSD_HEART, RMSSD_SOUND)

# ----- BETWEEN-GROUP t-tests (independent) -----
t_mean_heart <- t.test(Mean_HEART ~ group, data = df_wide)
t_mean_sound <- t.test(Mean_SOUND ~ group, data = df_wide)
w_rmssd_heart <- wilcox.test(RMSSD_HEART ~ group, data = df_wide)
w_rmssd_sound <- wilcox.test(RMSSD_SOUND ~ group, data = df_wide)

# ----- WITHIN-GROUP t-tests (paired) -----
# Mean comparison within FND
t_paired_mean_fnd <- t.test(
  df_wide$Mean_HEART[df_wide$group == "FND"],
  df_wide$Mean_SOUND[df_wide$group == "FND"],
  paired = TRUE)

# Mean comparison within HC
t_paired_mean_hc <- t.test(
  df_wide$Mean_HEART[df_wide$group == "HC"],
  df_wide$Mean_SOUND[df_wide$group == "HC"],
  paired = TRUE)

# RMSSD comparison within FND
w_paired_rmssd_fnd <- wilcox.test(
  df_wide$RMSSD_HEART[df_wide$group == "FND"],
  df_wide$RMSSD_SOUND[df_wide$group == "FND"],
  paired = TRUE)

# RMSSD comparison within HC
w_paired_rmssd_hc <- wilcox.test(
  df_wide$RMSSD_HEART[df_wide$group == "HC"],
  df_wide$RMSSD_SOUND[df_wide$group == "HC"],
  paired = TRUE)

# ----- Collect raw p-values -----
raw_pvals <- c(
  t_mean_heart$p.value,
  t_mean_sound$p.value,
  w_rmssd_heart$p.value,
  w_rmssd_sound$p.value,
  t_paired_mean_fnd$p.value,
  t_paired_mean_hc$p.value,
  w_paired_rmssd_fnd$p.value,
  w_paired_rmssd_hc$p.value)

# ----- Apply FDR correction -----
pvals_fdr <- p.adjust(raw_pvals, method = "fdr")

# ----- Print FDR-corrected results -----
labels <- c(
  "Mean_HEART (between-group)",
  "Mean_SOUND (between-group)",
  "RMSSD_HEART (between-group)",
  "RMSSD_SOUND (between-group)",
  "Mean (FND within-group)",
  "Mean (HC within-group)",
  "RMSSD (FND within-group)",
  "RMSSD (HC within-group)")

cat("## FDR-CORRECTED P-VALUES\n")
for (i in seq_along(labels)) {
  cat(labels[i], ": p (FDR) =", pvals_fdr[i], "\n")}

### specific values
mean(df_wide$Mean_HEART[df_wide$group == "FND"]) # 840ms = mean r-r interval for FND in focus on the heart
mean(df_wide$Mean_SOUND[df_wide$group == "FND"]) # 835ms = mean r-r interval for FND in focus on the sound
mean(df_wide$Mean_HEART[df_wide$group == "HC"]) # 921ms = mean r-r interval for FND in focus on the heart
mean(df_wide$Mean_SOUND[df_wide$group == "HC"])# 920ms = mean r-r interval for FND in focus on the heart

### specific values
mean(df_wide$RMSSD_HEART[df_wide$group == "FND"]) #  = mean HRV for FND in focus on the heart
mean(df_wide$RMSSD_SOUND[df_wide$group == "FND"]) #  = mean HRV for FND in focus on the sound
mean(df_wide$RMSSD_HEART[df_wide$group == "HC"]) #  = mean HRV for FND in focus on the heart
mean(df_wide$RMSSD_SOUND[df_wide$group == "HC"])#  = mean HRV for FND in focus on the heart
```
